### 2021-07-16 paper/main.tex
Quickly rattle off a draft before vacation

\documentclass[runningheads]{llncs}

\usepackage{amssymb,amsmath,mathtools,empheq,fancybox}

\usepackage{paralist}
\usepackage{url}
\usepackage{color}

\usepackage{textcomp,listings}
\usepackage{biblatex}
\usepackage{mymacros}
\usepackage{mathpartir}
\addbibresource{../bibliography.bib}

\newif\ifcomments
\commentstrue

\newcommand{\pr}[1]{{\color{red}``PR: #1''}}
\newcommand{\as}[1]{{\color{blue}``AS: #1''}}

\newif\ifoutline
\outlinetrue

\newcommand{\contents}[1]{\ifoutline{\color{blue}
    \begin{itemize}
    #1
    \end{itemize}
  }\fi}

\allowdisplaybreaks[1]

\title{Generalised Parikh Images of Regular Languages}
\author{some cool authors}
\institute{Uppsala University, Sweden}

\begin{document}
\maketitle

\begin{abstract}
  The image of the Parikh map is important in automata theory, offering a compact characterisation of an automaton. We contribute a novel understanding of how Parikh maps can be combined with arbitrary commutative monoid morphisms to efficiently represent a wide range of logics on automata and automata-like structures. Cases studied as examples include epistemic logic and string-length constraints in a string constraint solver. Furthermore, we show how this formulation can be efficiently implemented as a calculus in a theorem prover for succinct formulations of \Fudge{several constraints on strings}. In particular, our calculus is versatile enough to efficiently compute the Parikh image of a product of two automata, allowing our solver, \Ostrich{}, to solve \Fudge{X new instances} where it was previously constrained by the memory required to materialise the product. Finally, we show that this implementation in addition to being \Fudge{Z better} also offers \Fudge{X performance} improvements on \Fudge{Y real-world instances}, and in particular \Fudge{yields no regressions} in performance \Fudge{and, in fact, cures cancer, brings about world peace, and ends global hunger.}
\end{abstract}

\section{Introduction}

\contents{
\item key challenges: (1) complicated QE problem that was part of
  previous definitions; (2) languages with large alphabets described
  by symbolic automata; (3) regular languages described by products
  of automata
}

\subsection{Related Work}

\section{Overview}

\subsection{Generalised Parikh Images}\label{sec:generalised}

Formally, the \textit{Parikh map} over a context-free language $\Sigma = \left\{a_1, \ldots, a_k \right\}$ is defined as in \cite{kozen}:

$$
\begin{aligned}
& \psi: \Sigma^* \rightarrow \mathbb{N}^k \\
& \psi(s) = \left[\#a_1(s), \#a_2(s), \ldots, \#a_k(s)\right]
\end{aligned}
$$

That is, $\psi(s)$ is a vector of the number of occurrences of each character in the language for a given string $s$. For example, for  $\Sigma = \left \{ a, b\right\}$, we would have $\psi(abb) = \left[1, 2\right]$.

We define the image of this map, the \textit{Parikh image}, of some subset of the language $A \subseteq \Sigma^*$ as:

$$
\psi(A) = \left\{ \psi(x) | x \in A \right\}
$$

Thus we would have $\psi(\left\{ab, abb\right\}) = \left\{\left[1, 1\right], \left[1, 2\right]\right\}$.

\subsubsection{The Parikh image of a regular language expressed in Presburger arithmetic.}
Following~\cite{generate-parikh-image}, we define the Parikh Image of a regular language recognised by a DFA $\mathcal{A} =  \langle Q, \Sigma,\delta, I, F \rangle$ as:

$$
\begin{aligned}
\psi(\mathcal{A}) := 
& \bigwedge_{\alpha \in \Sigma}
c_\alpha = \sum_{\delta \in \delta} t_\delta  
\text{ where $\alpha \in \delta$}\\
&\bigwedge_{q \in Q} \left (\text{$1$ if $q \in I$} +
\sum_{\delta = q' \xrightarrow{} q} t_\delta 
- \sum_{\delta = q\xrightarrow{}q'} t_\delta \right)
\begin{cases}
\geq 0 & \text{if $q \in I$} \\
= 0 & \text{ otherwise}
\end{cases}\\
& \bigwedge_{\delta = q \xrightarrow{} q'} t_\delta > 0 
\implies z_{'q} > 0 \\
& \bigwedge_{q, q' \in Q} z_{q'} > 0 
\implies 
\begin{cases}
\bigvee\limits_{\delta = q \xrightarrow{} q'} z_{q'} = z_{q} + 1 \land t_\delta > 0 \land z_{q} > 0 & \text{if $q \not\in  I$} \\
z_{q'} = 1 \land t_\delta > 0& \text{if $q \in I$}
\end{cases}
\end{aligned}
$$

where all variables $z_i, t_i$ are existentially quantified and the variables $c_\alpha$ make up the actual image.

\subsubsection{Generalising the Parikh image}

Another way of viewing the Parikh map is as a monoid homomorphism $p:\: \left(\Sigma^*, \cdot, \epsilon \right) \to (\mathbb{Z}^\Sigma, +, \vec{0})$, where $\cdot$ is the string concatenation operation, the objects of the right-hand-side monoid are character counts, and $+$ is standard vector addition. Note that while the left monoid does not commute, the right one does.

This viewpoint enables us to generalise the Parikh map and its image further to arbitrary monoid morphisms $h:\: \Sigma^* \to M$ where $M$ is a commutative monoid. It then follows from the universal mapping property that any such morphism $h$ can also be expressed in terms of the Parikh map, as $h' \circ p$. A useful example of such a morphism might be computing the length of a string, which could easily be recast in terms of the Parikh map by summing the individual character counts of the vector.

\contents{
\item Examples
\item Fix the formalised version
\item Verify that the category theory is correct
}

\subsection{Lazy Computation of Parikh Images for Regular Languages}

\contents{
\item Lazy expansion of the Parikh conditions for a symbolic automaton
\item Finding elements vs.\ computing the complete Parikh image
\item Lazy product computation
}

\section{A Calculus of Parikh Images}

\contents{
\item Preliminaries, the underlying calculus, what are rules
\item Predicates used to represent Parikh images
\item Our calculus rules
\item Statement of properties, correctness, complexity
}

\section{Parikh Images from Products of Automata}

\contents{
\item additional rules needed for products
\item backjumping and conflict-driven learning
}

We assume an NFA $\Automaton = \Tuple{\States, \Alphabet, \Transitions, \Initial,
\Accepting}$ with $\NrTransitions$ transitions $\Transitions =
\Set{\EllipsisSequence{\Transitions}{\NrTransitions}}$ where we describe each
such transition $\Transitions_i$ from node $q$ to node $q'$ with label $\Label \in \Alphabet$ as
$\Transitions_i = \FromLabelTo{q}{\Label}{q'}$.

Treating $\Automaton$ as a graph with vertices $Q$ and edges $\delta$, we use
the term \textit{separating cut} of a set of \textit{transitions} $T =
\Set{\Transitions_i, \ldots, \Transitions_n}, \SeparatingCut(T)$ to refer to any set
of transitions whose removal causes $T$ to be unreachable from any state in
$\Initial$, with the meaning that a transition is reachable iff its starting
node is. Note that if $T$ contains a transition $e =
\FromLabelTo{v}{\Label}{v'}$ such that $v \in \Initial$, $\SeparatingCut(T) =
\emptyset$.

We will follow the notation of~\cite{generate-parikh-image} and simultaneously talk
about $\Automaton$ as a graph and an automaton. Moreover, we will continuously
refer to the subgraph produced by keeping only the transitions/edges whose
corresponding variables in $\TransitionVec$ are positive ($> 0$). An edge will
be called \textit{selected} if it is in this subgraph. An edge that is known to be zero
will conversely be called \textit{deselected}. An edge whose corrsponding term has no
known status is called unknown. Formally, we define these as follows:

\begin{definition}
  $\Selected(x)(\SomeClause)$ means that $x > a$ for some $a \geq 0$ is a clause
  in $\SomeClause$.
  \end{definition}

  \begin{definition}
  $\Deselected(x)(\SomeClause)$ means that $x = 0$ or $x < a$ for some $a \leq 1$
  is a clause in $\SomeClause$.
  \end{definition}

  \begin{definition}
  $\Unknown(x)(\SomeClause) = \lnot\Selected(x)(\SomeClause)\land\lnot\Deselected(x)(\SomeClause)$.
  \end{definition}

  Additionally, we define a helper predicate:

  \begin{definition}
    $\Connected(\Automaton)(\SomeClause)$ holds when every $\Selected$ or $\Unknown$ transition in $\Automaton$ is reachable from its initial state.
  \end{definition}

  With these preliminaries out of the way, we can define our main predicate:

\begin{definition}
$\PredicateInstance$ is true exactly when:
\begin{itemize}
\item $h:\: \Alphabet^* \rightarrow \SomeMonoid$, a morphism to a commutative product monoid $\SomeMonoid = \prod_{i} \SomeMonoid_i$.
\item $\MonoidElement$ is an element of $\SomeMonoid$.
\item $\TransitionVec$ is a vector of terms such that each term correspond to a
  transition in $\Automaton$.
\end{itemize}
\end{definition}

To simplify equations, we let $h(t)$ for some transition $t \in \Transitions$ to
mean the application of $h$ to $t$'s label. Similarly, we use the notation
$\TransitionVec_t$ to refer to transition $t \in \Transitions$'s corresponding
term, allowing $\TransitionVec$ to be directly indexed by the transitions
w.l.o.g from integer indices.

\begin{mathpar}
  \inferrule*[left=Propagate, right=\textnormal{$C = \SeparatingCut(T)$, $\exists t \in T \::\: \lnot \Deselected(t)$}]
    {\AndComp
      {\Transitions_t \in T}
      {\TransitionVec_{\Transitions_t} = 0} 
    \land \PredicateInstance \land \\
    \AndComp{\Transitions_c \in C}{\TransitionVec_{\Transitions_c} = 0} \land \SomeClause}
    {\PredicateInstance \land \AndComp{\Transitions_c \in C}{\TransitionVec_{\Transitions_c} = 0} \land \SomeClause}

  \inferrule*[left=Expand, right=\textnormal{Precisely once}]
    {\FlowEq(\Automaton) \land \Connected(\Automaton )\: \land \\ 
    y = \sum_{t \in \Transitions}{\TransitionVec_t \MonoidProduct h(t)} \land \AndComp{i \in 1,\ldots,t}{\TransitionVec_i \geq 0} \land \PredicateInstance \land \SomeClause}
    {\PredicateInstance \land \SomeClause}

  \inferrule*[left=Split, right=\textnormal{if $\Unknown(\TransitionVec_i)(\SomeClause)$}]
  {\TransitionVec_i = 0 \land \PredicateInstance \land \SomeClause \\ | 
  \\ \TransitionVec_i > 0 \land \PredicateInstance \land \SomeClause}
    {\PredicateInstance{} \land \SomeClause}

\inferrule*[left=Subsume-Main, right=\textnormal{if no instances of $\Connected(\Automaton)$ in $\SomeClause$}]
  {\SomeClause}
  {\PredicateInstance{} \land \SomeClause}

  \inferrule*[left=Subsume-Connected, right=\textnormal{if \KnownConnected}]
  {\SomeClause}
  {\Connected({\Automaton}) \land \SomeClause}
\end{mathpar}
where $\FlowEq(\Automaton)$ are the flow-balancing part of the Parikh image from Section~\ref{sec:generalised} for an automaton $\Automaton = \Tuple{\States, \Alphabet, \Transitions, \Initial,\Accepting}$:

$$
\begin{aligned}
& \In(q) = \text{$1$ if $q \in I$} + \sum_{i \in 1,\ldots,t \SuchThat \Transitions_i = \FromLabelTo{*}{}{q}} \TransitionVec_i\\
& \Out(q) = \sum_{i \in 1,\ldots,t \SuchThat \Transitions_i = \FromLabelTo{q}{}{*}} \TransitionVec_i\\
& \FlowEq = \AndComp{q \in Q}{\In(q) - \Out(q)}
\text{ $\geq 0$ if $q \in F$, $= 0$ otherwise}
\end{aligned}
$$

and $\KnownConnected$ corresponds to the following, implying guaranteed
connectedness (or, conversely, the non-existence of cuts disagreeing with $\TransitionVec$):
$$
\forall{C, T}\: C = \SeparatingCut(T) \implies \forall{i} \: \Transitions_i \in T \land \TransitionVec_i > 0 \implies \forall{j} \: \Transitions_j \in C \implies \TransitionVec_j = 0
$$

Additionally, we assume the existence of a rule \PresburgerClose{},
corresponding to a sound and complete solver for Presburger formulae.

The $\Propagate{}$ rule allows us to propagate connectedness across
$\Automaton$. It states that we are only allowed to "use" transitions attached
to a reachable state, and is necessary to ensure connectedness in the presence
of cycles. \textsc{Expand} expands the predicate into its most basic rules; one
set of linear equations connecting $\TransitionVec$ and $\MonoidElement$,
and the linear flow equations of the standard Parikh image formulation.

Finally, $\Split{}$ allows us to branch on the proof tree by first trying to
exclude a contested edge from a potential solution and then concluding that it
must be included.

A decision procedure for our predicate in a tableau-based automated theorem
prover would start by expanding the predicate using the $\Expand{}$ rule. For
many instances of the predicate, this would be enough to induce subsumption; as
as long as the DFA contains no loops that could be disconnected from a minimum
spanning tree (MST) of the automaton.

\subsection{Generalising to products of automata}

To generalise the calculus to calculations on products of automata, we change the main predicate to arbitrarily many automata:
\begin{definition}
  $\ProductPredicateInstance$ is true exactly when:
  \begin{itemize}
  \item $h:\: \Alphabet^* \rightarrow \SomeMonoid$, a morphism to a commutative product monoid $\SomeMonoid = \prod_{i} \SomeMonoid_i$.
  \item $\MonoidElement$ is a vector of elements of $\SomeMonoid$.
  \item $\Automaton = \prod_{i} \Automaton_i$
  \item each $\TransitionVec_i$ is a vector of terms such that each term correspond to a
    transition in $\Automaton_i$.
  \item $y = \sum_{t \in \Transitions} h(t)$
  \end{itemize}
  \end{definition}

  % FIXME y is a vector, and the plus of the sum is on the commutative monoid
  % FIXME backport this to the previous predicate's definition

  First we extend $\Expand$ to generate flow equations and instances of $\Connected$ for each automaton:
  \begin{mathpar}
    \inferrule*[left=Expand]
      {\AndComp{i}{\FlowEq(\Automaton_i) \land \Connected(\Automaton_i)}\ldots}
      {\ProductPredicateInstance \land \SomeClause}
  \end{mathpar}

\contents{
\item the rule MATERIALISE allows us to materialise a new partial product
\item SUBSUME now terminates iff we have fully materialised
}

\section{Implementation and Experiments}

\contents{
\item length constraints
\item Parikh automata, automata with registers
\item Model-checking examples
}

\section{Conclusions}

\clearpage
%\bibliographystyle{splncs03}
\printbibliography

\end{document}