### 2021-10-06 abstract/main.tex
Add abstract

\documentclass[runningheads]{llncs}

\usepackage{amssymb,amsmath,mathtools,empheq,fancybox}

\usepackage{paralist}
\usepackage{url}
\usepackage{color}

\usepackage{textcomp,listings}
\usepackage{biblatex}
\usepackage{mymacros}
\usepackage{mathpartir}
\addbibresource{../bibliography.bib}

\newif\ifcomments
\commentstrue

\newcommand{\pr}[1]{{\color{red}``PR: #1''}}
\newcommand{\as}[1]{{\color{blue}``AS: #1''}}

\newif\ifoutline
\outlinetrue

\newcommand{\contents}[1]{\ifoutline{\color{blue}
    \begin{itemize}
    #1
    \end{itemize}
  }\fi}

\allowdisplaybreaks[1]

\title{A Solver for Generalised Parikh Images of Conjunctions of Regular Languages}
\author{some cool authors}
\institute{Uppsala University, Sweden}

\begin{document}
\maketitle

\section{Introduction}

Parikh maps and their image appears naturally as part of many operations in
model checking and the solving of string constraints in automata-based string
solvers such as \Ostrich{} \cite{ostrich}. While it is possible to compute the Parikh image of
an automaton using the method described in \cite{generate-parikh-image}, this
method produces large existentially quantified clauses which are costly to
eliminate and in practice make many real-world problems intractable.
Furthermore, many modern applications require automata with symbolic labels to
handle large alphabets, which has no immediate correspondence in the methods described. Automata-based string solvers like \Ostrich{} also compute
Parikh images on products of automata, derived from conjunctions of string
constraints. Under previous methods this would require the computation of the
product before the Parikh image can be computed, running the risk of an
exponential blow-up.

Addressing these concerns, our approach allows us to extend the computation of
Parikh images to handle symbolic transition labels naturally, while also
allowing us to interleave the computation of arbitrarily deep products of
automata with the Parikh image of their product. This allows us to let both
calculations inform each other, allowing the computation of partial products to
inform the computation of the product and vice versa. Moreover, the scheme
allows us to learn interesting facts (implied clauses) about the problem.

Based on these insights, we present a tool to solve linear constraints on
symbolic automata with counters, amounting to constraints on the Parikh image of
their product. We are able to generate symbolic as well as concrete Parikh images.

\section{Background}

Formally, the \textit{Parikh map} over a context-free language $\Sigma = \left\{a_1, \ldots, a_k \right\}$ is defined as in \cite{kozen}:

$$
\begin{aligned}
& \psi: \Sigma^* \rightarrow \mathbb{N}^k \\
& \psi(s) = \left[\#a_1(s), \#a_2(s), \ldots, \#a_k(s)\right]
\end{aligned}
$$

That is, $\psi(s)$ is a vector of the number of occurrences of each character in the language for a given string $s$. For example, for  $\Sigma = \left \{ a, b\right\}$, we would have $\psi(abb) = \left[1, 2\right]$.

We define the image of this map, the \textit{Parikh image}, of some subset of the language $A \subseteq \Sigma^*$ as:

$$
\psi(A) = \left\{ \psi(x) | x \in A \right\}
$$

Thus we would have $\psi(\left\{ab, abb\right\}) = \left\{\left[1, 1\right], \left[1, 2\right]\right\}$.

\subsection{Generalised Parikh Images}\label{sec:generalised}

Another way of viewing the Parikh map is as a monoid homomorphism $p:\: \left(\Sigma^*, \cdot, \epsilon \right) \to (\mathbb{Z}^\Sigma, +, \vec{0})$, where $\cdot$ is the string concatenation operation, the objects of the right-hand-side monoid are character counts, and $+$ is standard vector addition. Note that while the left monoid does not commute, the right one does.

This viewpoint enables us to generalise the Parikh map and its image further to arbitrary monoid morphisms $h:\: \Sigma^* \to M$ where $M$ is a commutative monoid. It then follows from the universal mapping property that any such morphism $h$ can also be expressed in terms of the Parikh map, as $h' \circ p$.

A useful example of such a morphism might be computing the length of a string,
which could be recast in terms of the Parikh map by summing the individual
character counts of the vector: $h':\: (\mathbb{Z}^\Sigma, +, \vec{0}) \to
(\mathbb{Z}, +, 0) = \vec{x} \to \sum_{i \in \Sigma} x_i$, where the respective
operators $+$ is element-wise vector addition and integer addition respectively.
This allows us to apply selective logics, symbolic logics, and arithmetic on
counter variabels to the Parikh image calculus.

\section{Lazy Computation of Parikh Images for Regular Languages}

%\bibliographystyle{splncs03}
\printbibliography

\end{document}