### 2022-04-04 paper/main.tex
Today's edits

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=true}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2023}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\usepackage{amsthm}
\usepackage{amsmath,mathtools,empheq,fancybox}
\usepackage{paralist}
\usepackage{url}
\usepackage{color}
\usepackage{textcomp,listings}
\usepackage{array}
\usepackage{mymacros}

\usepackage{mathpartir}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\newif\ifcomments
\commentstrue
\newif\ifoutline
\outlinetrue

\newcommand{\contents}[1]{\ifoutline{\color{blue}
    \begin{itemize}
    #1
    \end{itemize}
  }\fi}

\allowdisplaybreaks[1]

\begin{document}

%% Title information
\title{An Efficient Calculus for Generalised Parikh Images of Regular Languages}
                                        %% when present, will be used in
                                        %% header instead of Full Title.
\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{First1 Last1}
\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position1}
  \department{Department1}              %% \department is recommended
  \institution{Institution1}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{City1}
  \state{State1}
  \postcode{Post-Code1}
  \country{Country1}                    %% \country is recommended
}
\email{first1.last1@inst1.edu}          %% \email is recommended

%% Author with two affiliations and emails.
\author{First2 Last2}
\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position2a}
  \department{Department2a}             %% \department is recommended
  \institution{Institution2a}           %% \institution is required
  \streetaddress{Street2a Address2a}
  \city{City2a}
  \state{State2a}
  \postcode{Post-Code2a}
  \country{Country2a}                   %% \country is recommended
}
\email{first2.last2@inst2a.com}         %% \email is recommended
\affiliation{
  \position{Position2b}
  \department{Department2b}             %% \department is recommended
  \institution{Institution2b}           %% \institution is required
  \streetaddress{Street3b Address2b}
  \city{City2b}
  \state{State2b}
  \postcode{Post-Code2b}
  \country{Country2b}                   %% \country is recommended
}
\email{first2.last2@inst2b.org}         %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  The image of the Parikh map is important in automata theory, offering a
  compact characterisation of an automaton. We contribute a novel understanding
  of how Parikh maps can be combined with arbitrary commutative monoid morphisms
  to efficiently represent a wide range of logics on automata and automata-like
  structures. This allows us to \Fudge{ensure world peace forever and express
  loop invariants that solve the halting problem for any practical program}.
  Furthermore, we show how this formulation can be efficiently implemented as a
  calculus in a theorem prover supporting Presburger logic. In particular, our
  calculus allows us to solve the common problem in string solvers of computing
  the Parikh image of products of arbitrarily many automata. We demonstrate this
  by implementing a tool called \Catra{}, which we use to solve a set of
  constraints produced by \Fudge{the PyEx benchmarks} when solved by the
  \Ostrich{} string constraint solver. We show that this implementation in
  addition to \Fudge{having a lower memory footprint} than the standard eager
  approach executed on the same underlying solver also \Fudge{strictly}
  outperforms the nuxmv model checker on the same problem\Fudge{, as well as the
  over-approximation recently described by~\citeauthor{approximate-parikh} and
  the original approach described by \citeauthor{generate-parikh-image}.}

\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}

Parikh maps and their image appears naturally as part of many operations in
model checking and the solving of string constraints in automata-based string
solvers such as \Ostrich{}, notably in representing constraints on string
lengths. While it is possible to compute the Parikh image of any context-free
language using the method described in~\cite{generate-parikh-image}, this method
produces existentially quantified clauses and take at most linear time to
produce~\cite{muscholl-linear}. However, the resulting existentially quantified
clauses are costly to eliminate, in practice making many real-world problems
intractable. Furthermore, applications to string analysis in theorem provers
require symbolic labels to handle Unicode alphabets \Fudge{citation needed}.
Conjunctions of string constraints finally, and most crucially, lead to the
computation of Parikh images of products of regular languages. Applying the
approach in \cite{generate-parikh-image} would require the computation of the
product before the Parikh image can be computed. In several instances we have
observed, the materialisation of the product of the automata exhausts the memory
of any reasonable machine due to the exponential blow-up of the product.

Addressing these concerns, our specialised calculus for Parikh images of
products of regular languages allows us to handle symbolic transition labels
naturally, while also allowing us to interleave the computation of arbitrarily
deep products of automata with the Parikh image of their product. This allows us
to let both calculations inform each other, eliminating unnecessary work, and
pruning the size of the partial products used in the computation. Moreover, the
method can be used iteratively to, as it were, chew off bite-sized chunks of the
product, thereby avoiding the problem of memory-outs.

\subsection{Motivating Example}

\section{Overview}

We define a non-deterministic automaton~$\Automaton$ with alphabet~$\Alphabet$
as $\Automaton = \AutomatonTuple$ where $\Transitions = \States \times \Alphabet
\times \States$, $\States$ is its states, $\InitialState$ is w.l.o.g. assumed to
be the single initial state, and $\AcceptingStates$ is its set of accepting
states. Note that we will later generalise the labels of transitions further.

\subsection{The Parikh Map}
Formally, the \textit{Parikh map} over a context-free language $\Strings=
\left\{a_1, \ldots, a_k \right\}$ is defined as in \cite{kozen}:

$$
\begin{aligned}
& \ParikhMap: \MapFromTo{\Strings}{\natural^k} \\
& \ParikhMap(s) = \VectorLiteral{\#a_1(s), \#a_2(s), \ldots, \#a_k(s)}
\end{aligned}
$$

That is, $\ParikhMap(s)$ is a vector of the number of occurrences of each
character in the language for a given string $s$. For example, for  $\Strings =
\Set{a, b}$, we would have $\ParikhMap(abb) = \VectorLiteral{1, 2}$.

We define the image of this map, the \textit{Parikh image}, of some subset of
the language $\Language \subseteq \Strings$ as:

\[
\ParikhMap(\Language) = \Set{\ParikhMap(x) \SuchThat x \in \Language}
\]

Thus, we would have $\ParikhMap(\left\{ab, abb\right\}) = \left\{\left[1,
1\right], \left[1, 2\right]\right\}$.

Parikh's theorem states that ignoring order, any context-free grammar has a
letter-equivalent regular language (c.f.~\cite{construction} for a construction of
such automata from context-free grammars). However, there are languages that are
not context-free that also have semilinear images under~$\ParikhMap$ (e.g.
$\ParikhMap(\Set{a^nb^nc^n \SuchThat n \geq 0}) = \ParikhMap((abc)^*) = \#a =
\#b = \#c \land \#a \geq 0$).

\contents {
  \item parikh images of regular images are semilinear, therefore presburger
  \item presburger is semilinear
}

\subsection{The Parikh image of a regular language expressed in Presburger arithmetic}

Following~\cite{generate-parikh-image}, we define the Parikh Image of a regular
language recognised by a DFA $\Automaton = \AutomatonTuple$ as:

\[
\begin{aligned}
\psi(\mathcal{A}) := 
& \bigwedge_{\alpha \in \Sigma}
c_\alpha = \sum_{\delta \in \delta} t_\delta  
\text{ where $\alpha \in \delta$}\\
&\bigwedge_{q \in Q} \left (\text{$1$ if $q \in I$} +
\sum_{\delta = q' \xrightarrow{} q} t_\delta 
- \sum_{\delta = q\xrightarrow{}q'} t_\delta \right)
\begin{cases}
\geq 0 & \text{if $q \in I$} \\
= 0 & \text{ otherwise}
\end{cases}\\
& \bigwedge_{\delta = q \xrightarrow{} q'} t_\delta > 0 
\implies z_{'q} > 0 \\
& \bigwedge_{q, q' \in Q} z_{q'} > 0 
\implies 
\begin{cases}
\bigvee\limits_{\delta = q \xrightarrow{} q'} z_{q'} = z_{q} + 1 \land t_\delta > 0 \land z_{q} > 0 & \text{if $q \not\in  I$} \\
z_{q'} = 1 \land t_\delta > 0& \text{if $q \in I$}
\end{cases}
\end{aligned}
\]

where all variables $z_i, t_i$ are existentially quantified and the variables
$c_\alpha$ make up the actual image. Feeding this definition into a Presburger
solver and performing quantifier elimination on $z_i, t_i$ would produce the
Parikh image in its Presburger form.

In this paper we use this example (sometimes appropriately modified) as the
baseline approach. Notably, the approach presented in this paper can at its core
be described as lazily enforcing the connectedness constraint of this method
(the final two clauses) while also interleaving the (similarly lazy) computation
of products of automata.

\subsubsection{An Example}

\Fudge{AN EXAMPLE!}

\subsection{Generalised Parikh Images}\label{sec:generalised}

To generalise the logic over Parikh images, we introdouce a morphism $\Map:\: \Sigma
\to \Monoid$ where $\Monoid$ is a commutative monoid where the following holds:
\begin{itemize}
  \item $\Map(\epsilon) = 0$, where $0$ is the neutral element of $\Monoid$
  \item $\Map(a \Concat b) = \Map(a) + \Map(b)$
\end{itemize}

\contents{
  \item The Anca Muscholl paper?
  \item Describe how a "simpler" function can accelerate solving
}

\subsubsection{Example~1: String length}

A useful example of such a morphism can express string length. We let $\Monoid =
\left(\mathbb{Z};+;0\right)$ and let $\Map(x) = 1$. Then the length of a string $s
= s_1 \RepeatSum{\Concat}  s_n$ is given by $\Map(s) = \sum \Map(s_i) = 1
\RepeatSum{+} 1$. Similarly, vectors and vector addition can be used to obtain
the regular Parikh map.

\subsubsection{Example~2: Parikh Automata}

Another application for this generalisation is to express Parikh
automata~\cite{parikh-automata}, and this is indeed what the implementation
described in Section~\ref{sec:implementation} does.

\contents{
 \item what is parikh automata
 \item Give an actual example of a parikh automata
}

\subsection{Lazy Computation of Parikh Images for Regular Languages}

We assume an NFA $\Automaton = \AutomatonTuple$ with $\NrTransitions$
transitions $\Transitions =
\Set{\EllipsisSequence{\Transitions}{\NrTransitions}}$ where we describe each
such transition $\Transitions_i$ from state $\State$ to state $\State'$ with
label $\Label \in \Alphabet$ as $\Transitions_i =
\FromLabelTo{\State}{\Label}{\State'}$. When one or more of the states and
labels of a transition are uninteresting we will omit it.

For convenience, we introduce the following supporting concepts and notations:

\begin{definition}
  A \textit{path} $\Path = \PathEnumeration$ of an automaton~$\Automaton$ with
  transitions $\Transitions$ represents a path through $\Automaton$ using
  transitions in $\Transitions$ (i.e. $\Transitions(\State_k, \Label_{k+1},
  \State_{k+1})$ holds for every $k$), passing zero or more labels $\Label_1,
  \ldots, \Label_n$. The path must begin in the initial state, i.e.~$\State_0 =
  \InitialState$. However, the ending state, $\State_n$, is not necessarily an
  accepting state.
  \end{definition}

\begin{definition}
  Moreover, we also talk about the \textit{set of paths} of an automaton,
  $\Paths(\Automaton)$, a possibly infinite (if $\Automaton$ has loops) set of
  valid paths through $\Automaton$. Additionally, we use the
  notation~$\Paths(\Automaton, \State)$ to mean all paths ending in
  state~$\State$.
\end{definition}

\begin{definition}
  The \textit{word} of a path $\WordOf(\Path) = \Label_1 \RepeatSum{\Concat} \Label_n$ is
  the word read out on its labels.
\end{definition}

\begin{definition}
  The \textit{states} of a path $\StatesOf(\Path) = \Set{\State \SuchThat
  \FromLabelTo{\State}{}{\State'} \in \Path \text{ or }
  \FromLabelTo{\State'}{}{\State} \in \Path}$ are the states visited along a
  path. Note that $\InitialState \in \StatesOf(\Path)$ for every path since all
  paths start in the initial state.
\end{definition}

\begin{definition}
  A \textit{cut}, $C$ of an automaton~$\Automaton = \AutomatonTuple$,
  $\SeparatingCut(C, \Automaton, \State)$, is a minimal set of transitions whose
  removal from~$\Transitions$ would cause~$\State \in \States$ to be unreachable
  from the initial state~$\InitialState$. $\SeparatingCut(\Automaton,
  \InitialState)$ is crucially undefined since it cannot be made unreachable
  from itself by definition.
\end{definition}

\begin{definition}
 The \textit{transition count}, $\TransitionCount(\Transition, \Path)$ is the
 number of times a transition $\Transition =
 \FromLabelTo{\State_1}{\Label}{\State_2} \in \Transitions$ appears in a path
 $p$.
\end{definition}

We then introduce the two predicates into our calculus with the following
definitions:

\begin{definition}
  The Parikh predicate, $\SinglePredicateInstance$, for some automaton
  $\Automaton = \AutomatonTuple$, modulo some map $\Map$ to a commutative monoid
  $\Monoid$ as described in Section~\ref{sec:generalised} and with a transition
  selection function $\Filter:\: \Transitions \to \Naturals$ holds when
  $\MonoidElement \in \Monoid$ is the Parikh image of $\Automaton$ modulo
  $\Map$, or more formally when there is an accepting path $\Path =
  \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ such that:
  \begin{itemize}
    \item $\Filter(\Transition) = \TransitionCount(\Transition, \Path)$
    \item $\MonoidElement = \Map(\WordOf(\Path))$
  \end{itemize}
\end{definition}

\begin{definition}
  $\Connected(\Automaton, \Filter)$ for some automaton $\Automaton =
  \AutomatonTuple$ holds when for every $\Transition =
  \FromLabelTo{\State}{}{\State'} \in \Transitions$, $\Filter(\Transition) > 0
  \implies \exists \Path \in \Paths(\Automaton)$ $\Filter(\Transition_i) > 0$
  for $\Transition_i \in p$, and $\State \in \StatesOf(p)$, or in words that
  there exists some $\Filter$-selected valid path that reaches $\Transition$'s
  starting state, $\State$. Intuitively, it represents the condition that
  $\Automaton$ is connected with respect to the selection function $\Filter$ for
  every transition. It is redundant to $\Image$ by design.
\end{definition}

Finally, we present the rules of our calculus for one automaton in Table~\ref{tbl:rules:single}. Note that we
operate on sets of symbols (terms, clauses). Additionally, we also use the
convention of splitting the formulas into linear inequalities ($\SomeInequalities$)
and any other clauses ($\SomeClause$).

We use the shorthand notation~$\Transitions_\Automaton$ to refer to the
transitions of an automaton~$\Automaton$. Additionally, for an
automaton~$\Automaton = \AutomatonTuple$ we allow mapping the selection function
like so: $\Filter(\Automaton) = \Tuple{\States, \InitialState, \AcceptingStates,
\Set{\Transition \in \Transitions \SuchThat \Filter(\Transition) > 0}}$, i.e.
$\Automaton$~with only the transitions for which~$\Filter$ is positive. In this
instance, the basis for the matched linear inequalities is
implicitly~$\SomeInequalities$.

These rules are meant to be executed in an automated theorem prover, and are
read bottom-up. Since a large part of our rules operate by adding and literally
matching linear (in)equalities in a proof goal, we use the shorthand of listing
the matched inequalities as antecedents. An example of this can be seen in
the~\Propagate{} rule.

Please note that the filtering function~$\Filter$ is evaluated symbolically in
these rules, and can in practice be read as a symbolic function from transitions
to terms (e.g.~\texttt{t} or~\texttt{t+1}), not dissimilar to~$\FlowEq$ below.
In our implementation, $\Filter$~is a vector of fresh unknown constants with the
same size as~$\Transitions$.

% Use a table!

We assume that every rule only fires when it would add a new clause. For
example, this means that we cannot use~\Split{} to split on the same term twice.
This suggests a proof strategy where you \Propagate{} when you can, \Split{}
when you must, and \Subsume{} when neither is possible anymore.

\begin{table}[h]
\begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
  Name & Rule & Prerequisites\\
  \midrule

  % EXPAND
  \Expand & 
    \inferrule
  {\Connected(\Automaton, \Filter) \land \FlowEq(\Automaton, \Filter) \land \MonoidElement = \sum_{\Transition \in \Transitions_\Automaton} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
  {\SinglePredicateInstance, \SomeInequalities, \SomeClause} & 
  None \\

  % SPLIT
  \Split & 
  \inferrule{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0 \mid \Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) > 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
  if $\Transition \in \Transitions_\Automaton$ \\

  % PROPAGATE
  \Propagate &
  \inferrule{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause, \Filter(\Transition) = 0}{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause} &
  if $t = \FromLabelTo{\State}{}{\State'} \in \Transitions_\Automaton, \SeparatingCut(C, \Automaton, \State)$\\

  % SUBSUME
  \Subsume &
  \inferrule{\SomeInequalities,\SomeClause}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
  \Split{} and \Propagate{} cannot be applied \\
  \bottomrule
  \end{tabular}
  \caption{Derivation rules for one automaton.}\label{tbl:rules:single}
\end{table}

We use the symbolic function $\FlowEq(\Automaton, \Filter)$ that generates a set
of existentially quantified linear inequalities with the following definition:
$$
\begin{aligned}
  & \FlowEq(\Automaton, \Filter) = \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)}
\text{ $\geq 0$ if $\State \in \AcceptingStates$, $= 0$ otherwise}\\
  & \In(\State, \Filter) = \left(\text{$1$ if $\State = \InitialState$, otherwise $0$}\right) + \sum_{\Transition = \FromLabelTo{\State_0}{}{\State} \in \Transitions} \Filter(\Transition)\\
  & \Out(\State, \Filter) = \sum_{\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions} \Filter(\Transition)
\end{aligned}
$$
Where we transparently assign fresh variables to each state $\State \in \States$
and existentially quantify them in the whole clause.

Additionally, we assume the existence of a rule \PresburgerClose{},
corresponding to a sound and complete solver for Presburger formulae, modulo the
monoid~$\Monoid$.

The $\Propagate{}$ rule allows us to propagate (dis-)connectedness across
$\Automaton$. It states that we are only allowed to use transitions attached to
a reachable state, and is necessary to ensure connectedness in the presence of
cycles in~$\Automaton$.

\textsc{Expand} expands the predicate into its most basic rules; one set of
linear equations synchronising the transitions mentioned by~$\Filter$ to the
corresponding Monoid element~$\MonoidElement$, and the linear flow equations of
the standard Parikh image formulation, as described by~\FlowEq. Since
$\Connected$ and $\Image$ are partially redundant and the difference is covered
by~$\FlowEq$, we can remove the instance of~$\Image$ when applying~$\Expand$. In
this sense, we split the semantics of the $\Image$~predicate into its counting
aspect (covered by $\FlowEq$) and its connectedness aspect (covered by
$\Connected$).

Finally, $\Split{}$ allows us to branch the proof tree by trying to exclude a
contested transition from a potential solution before concluding that it must be
included. Intuitively, this is what guarantees our ability to make forward
progress by eliminating paths through~$\Automaton$.

A decision procedure for our predicate in a tableau-based automated theorem
prover would start by expanding the predicate using the $\Expand{}$~rule. A
modestly clever theorem prover would perform algebraic substitution on the
underlying constants of~$\Filter$, boiling them down to choices of branches,
which depend on one single variable, and loop transitions. This logic
corresponds to the placement of counters for optimally edge-profiling the CFG of
a program, making up a MST of the automaton~\cite{path-profiling}.

\subsection{An Example}

\subsection{Finding the Presburger representation of a Parikh Image}
\contents{
  \item lazy qe
}

\subsection{The calculus is correct}

To be as clear as possible about what it means for our calculus to be correct,
we use more exact definitions than the traditional soundness and completeness.

\subsubsection{The calculus terminates}
\subsubsection{The calculus allows no values outside the Parikh image}
\subsubsection{The calculus excludes no values in the Parikh image}

\section{Extending and generalising the calculus}

\section{Parikh Images from Products of Automata}

\contents{
    \item additional rules needed for products
    }

\section{Backjumping and Learning No-Goods}

% To generalise the calculus to calculations on products of automata, we change the main predicate take arbitrarily many automata:
% \begin{definition}
%   $\ProductPredicateInstance$ is true exactly when:
%   \begin{itemize}
%   \item $h:\: \Alphabet^* \rightarrow \SomeMonoid$, a morphism to a commutative product monoid $\SomeMonoid = \prod_{i} \SomeMonoid_i$.
%   \item $\MonoidElement$ is a vector of elements of $\SomeMonoid$.
%   \item $\AutomatonVector$ is a vector of automata.
%   \item $\Automaton = \prod_{i} \AutomatonVector_i$
%   \item each $\TransitionVec_i$ is a vector of terms such that each term correspond to a
%     transition in $\AutomatonVector_i$, and $\TransitionVectors$ is a vector of those vectors. We use the terminology $\TransitionTerm(t)$ to refer to the term of the transition vector for a given transition $t$.
%   \item $y = \sum_{t \in \Transitions} h(t)$
%   \item $\SourceTransitions(\Automaton, t)$ is the set of transitions from the
%   term automata used to produce the transition $t$ in the product automaton
%   $\Automaton$.
%   \end{itemize}
%   \end{definition}

%   First we extend $\Expand$ to generate flow equations and instances of $\Connected$ for each automaton:
%   \begin{mathpar}
%     \inferrule*[left=Expand]
%       {\AndComp{\Automaton_i \in \AutomatonVector}{\FlowEq(\Automaton_i) \land \Connected(\Automaton_i)}\ldots}
%       {\ProductPredicateInstance \land \SomeClause}
%   \end{mathpar}

%   %  x(t) = sum(e : termProductEdges(t, default=0))
% Then we introduce the rule $\Materialise$, used to compute a partial product between two terms $\Automaton_i, \Automaton_j$:
% \begin{mathpar}
%   \inferrule*[left=Materialise, 
%   right=\textnormal{where $\Automaton_{k} = \Automaton_i \times \Automaton_j, \AutomatonVector' = (\AutomatonVector \cup \{  \Automaton_k \}) \setminus {\Automaton_i, \Automaton_j}$}]
%     {\AndComp{t \in \delta(\Automaton_{k})}{\TransitionTerm(t) = \sum_{t' \in \SourceTransitions(t, \Automaton_k)} t'}\\
%     \FlowEq(\Automaton_{k}) \land \Connected(\Automaton_{k}) \land
%     \\ \MonoidMap(\AutomatonVector', \TransitionVectors', \MonoidElement) \land \SomeClause} {\MonoidMap(\AutomatonVector, \TransitionVectors, \MonoidElement) \land \SomeClause}
% \end{mathpar}

% Note that we implicitly existentially quantify the fresh transition terms $\TransitionVec_{k+1}$ for the product.

% $\Materialise$ comes with its companion rule $\RemoveTermConnected$ that simply
% states we are allowed to remove any instances of $\Connected$ from terms that
% have been used in a product, and whose connectedness we therefore no longer care about:
% \begin{mathpar}
%   \inferrule*[left=$\RemoveTermConnected$, right=\textnormal{If $\Automaton_a$ not in $\Automaton_1\ldots\Automaton_k$}]
%     {\ProductPredicateInstance \land \SomeClause}
%     {\Connected(\Automaton_a) \land \ProductPredicateInstance \land \SomeClause}
% \end{mathpar}

% Finally, we add a side condition to $\Subsume$ to only allow subsumption for
% instances of $\MonoidMap$ with just one automaton, ensuring that we do not bail
% out before having computed the entire product.

\subsection{An Example}

\section{Implementation and Experiments}

\contents{
\item length constraints
\item Parikh automata, automata with registers
\item Model-checking examples e.g. \cite{epistemic-logic}
}

\section{Conclusions}

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

%% Bibliography
\bibliography{bibliography}

%% Appendix
\appendix
\section{Appendix}

Text of appendix \ldots

\end{document}