### 2022-05-03 paper/main.tex
More structure

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=true}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2023}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\usepackage{amsthm}
\usepackage{amsmath,mathtools,empheq,fancybox}
\usepackage{paralist}
\usepackage{url}
\usepackage{color}
\usepackage{textcomp,listings}
\usepackage{array}
\usepackage{mymacros}
\usepackage{microtype}
\usepackage{listings}
\usepackage{csquotes}

\usepackage{mathpartir}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\newif\ifcomments
\commentstrue
\newif\ifoutline
\outlinetrue

\newcommand{\contents}[1]{\ifoutline{\color{blue}
    \begin{itemize}
    #1
    \end{itemize}
  }\fi}

\allowdisplaybreaks[1]

\begin{document}

%% Title information
\title{An Efficient Calculus for Generalised Parikh Images of Regular Languages}
                                        %% when present, will be used in
                                        %% header instead of Full Title.
\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{First1 Last1}
\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position1}
  \department{Department1}              %% \department is recommended
  \institution{Institution1}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{City1}
  \state{State1}
  \postcode{Post-Code1}
  \country{Country1}                    %% \country is recommended
}
\email{first1.last1@inst1.edu}          %% \email is recommended

%% Author with two affiliations and emails.
\author{First2 Last2}
\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position2a}
  \department{Department2a}             %% \department is recommended
  \institution{Institution2a}           %% \institution is required
  \streetaddress{Street2a Address2a}
  \city{City2a}
  \state{State2a}
  \postcode{Post-Code2a}
  \country{Country2a}                   %% \country is recommended
}
\email{first2.last2@inst2a.com}         %% \email is recommended
\affiliation{
  \position{Position2b}
  \department{Department2b}             %% \department is recommended
  \institution{Institution2b}           %% \institution is required
  \streetaddress{Street3b Address2b}
  \city{City2b}
  \state{State2b}
  \postcode{Post-Code2b}
  \country{Country2b}                   %% \country is recommended
}
\email{first2.last2@inst2b.org}         %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  The image of the Parikh map is important in automata theory, offering a
  compact characterisation of an automaton. We contribute \Fudge{a novel}
  understanding of how Parikh maps can be combined with arbitrary commutative
  monoid morphisms to efficiently represent a wide range of logics on automata
  and automata-like structures. This allows us to \Fudge{ensure world peace
  forever and express loop invariants that solve the halting problem for any
  practical program}. Furthermore, we show how this formulation can be
  efficiently implemented as a calculus in a theorem prover supporting
  Presburger logic. In particular, our calculus allows us to solve the common
  problem in string solvers of computing the Parikh image of products of
  arbitrarily many automata. We demonstrate this by implementing a tool called
  \Catra{}, which we use to solve a set of constraints produced by \Fudge{the
  PyEx benchmarks} when solved by the \Ostrich{} string constraint solver. We
  show that this implementation in addition to \Fudge{having a lower memory
  footprint} than the standard eager approach executed on the same underlying
  solver also \Fudge{strictly} outperforms the \Nuxmv{} model checker on the
  same problem\Fudge{, as well as both the over-approximation recently described
  by~\citeauthor{approximate-parikh} and the original approach described by
  \citeauthor{generate-parikh-image}.}

\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{parikh images, string solvers, model checking}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}

The Parikh map is a characterisation of a context-free language as a vector of
its character counts, famously a semilinear set~\cite{parikh-theorem}. It
appears naturally as part of many operations in model checking and the solving
of string constraints in automata-based string solvers such as \Ostrich{},
notably in representing constraints on string lengths.

It is possible to compute the Parikh image of any context-free language using
the method described in~\cite{generate-parikh-image}. Later improvements have
produced a construction taking at most linear time to
produce~\cite{muscholl-linear}. However, the resulting existentially quantified
clauses are costly to eliminate, in practice making many real-world problems
intractable.

In addition to costly elimination of the existentially quantified variables,
another problem arises for the common case of solving length constraints in
string solvers. Conjunctions of string constraints lead to the computation of
Parikh images of products of regular languages. Applying the approach
in~\cite{generate-parikh-image} would require the computation of the product
before its Parikh image. In several instances we have observed while solving
real-world string constraints, the computation the product of automata exhausts
the memory of any machine due to the exponential blow-up in size of the product,
quickly becoming intractable as the number of terms in the product increases.

Addressing these concerns, our specialised calculus for Parikh images of
products of regular languages allows us to interleave the computation of
arbitrarily deep products of automata with the product's Parikh image. This
allows us to let both calculations inform each other, eliminating unnecessary
work, and pruning the size of the partial products considered in the
computation. Moreover, the method can be used iteratively to tackle smaller
chunks of the product incrementally, thereby avoiding running out of memory.

\subsection{Motivating Example}

\section{Overview}

We define a non-deterministic automaton~$\Automaton$ with alphabet~$\Alphabet$
as $\Automaton = \AutomatonTuple$ where $\Transitions = \States \times \Alphabet
\times \States$, $\States$ is its states, $\InitialState$ is w.l.o.g. assumed to
be the single initial state, and $\AcceptingStates$ is its set of accepting
states. Note that we will later generalise the labels of transitions further.

\subsection{The Parikh Map}
Formally, the \textit{Parikh map} over a context-free language $\Strings=
\left\{a_1, \ldots, a_k \right\}$ is defined as in \cite{kozen}:

$$
\begin{aligned}
& \ParikhMap: \MapFromTo{\Strings}{\natural^k} \\
& \ParikhMap(s) = \VectorLiteral{\#a_1(s), \#a_2(s), \ldots, \#a_k(s)}
\end{aligned}
$$

That is, $\ParikhMap(s)$ is a vector of the number of occurrences of each
character in the language for a given string $s$. For example, for  $\Strings =
\Set{a, b}$, we would have $\ParikhMap(abb) = \VectorLiteral{1, 2}$.

We define the image of this map, the \textit{Parikh image}, of some subset of
the language $\Language \subseteq \Strings$ as:

\[
\ParikhMap(\Language) = \Set{\ParikhMap(x) \SuchThat x \in \Language}
\]

Thus, we would have $\ParikhMap(\left\{ab, abb\right\}) = \left\{\left[1,
1\right], \left[1, 2\right]\right\}$.

Parikh's theorem states that ignoring order, any context-free grammar has a
letter-equivalent regular language (c.f.~\cite{construction} for a construction of
such automata from context-free grammars and~\cite{bounds} for bounds on its size). However, there are languages that are
not context-free that also have semilinear images under~$\ParikhMap$ (e.g.
$\ParikhMap(\Set{a^nb^nc^n \SuchThat n \geq 0}) = \ParikhMap((abc)^*) = \#a =
\#b = \#c \land \#a \geq 0$).

\contents {
  \item parikh images of regular images are semilinear, therefore presburger
  \item presburger is semilinear
}

\subsection{The Parikh image of a regular language expressed in Presburger arithmetic}

Following~\cite{generate-parikh-image}, we define the Parikh Image of a regular
language recognised by a DFA $\Automaton = \AutomatonTuple$ as:

\[
\begin{aligned}
\psi(\mathcal{A}) := 
& \bigwedge_{\alpha \in \Sigma}
c_\alpha = \sum_{\delta \in \delta} t_\delta  
\text{ where $\alpha \in \delta$}\\
&\bigwedge_{q \in Q} \left (\text{$1$ if $q \in I$} +
\sum_{\delta = q' \xrightarrow{} q} t_\delta 
- \sum_{\delta = q\xrightarrow{}q'} t_\delta \right)
\begin{cases}
\geq 0 & \text{if $q \in I$} \\
= 0 & \text{ otherwise}
\end{cases}\\
& \bigwedge_{\delta = q \xrightarrow{} q'} t_\delta > 0 
\implies z_{'q} > 0 \\
& \bigwedge_{q, q' \in Q} z_{q'} > 0 
\implies 
\begin{cases}
\bigvee\limits_{\delta = q \xrightarrow{} q'} z_{q'} = z_{q} + 1 \land t_\delta > 0 \land z_{q} > 0 & \text{if $q \not\in  I$} \\
z_{q'} = 1 \land t_\delta > 0& \text{if $q \in I$}
\end{cases}
\end{aligned}
\]

where all variables $z_i, t_i$ are existentially quantified and the variables
$c_\alpha$ make up the actual image. Feeding this definition into a Presburger
solver and performing quantifier elimination on $z_i, t_i$ would produce the
Parikh image in its Presburger form.

In this paper we use this example (sometimes appropriately modified) as the
baseline approach. Notably, the approach presented in this paper can at its core
be described as lazily enforcing the connectedness constraint of this method
(the final two clauses) while also interleaving the (similarly lazy) computation
of products of automata.

\subsubsection{An Example}

\Fudge{AN EXAMPLE!}

\subsection{Generalised Parikh Images}\label{sec:generalised}

To generalise the logic over Parikh images, we introdouce a morphism $\Map:\: \Sigma
\to \Monoid$ where $\Monoid$ is a commutative monoid where the following holds:
\begin{itemize}
  \item $\Map(\epsilon) = 0$, where $0$ is the neutral element of $\Monoid$
  \item $\Map(a \Concat b) = \Map(a) + \Map(b)$
\end{itemize}

\contents{
  \item The Anca Muscholl paper?
  \item Describe how a "simpler" function can accelerate solving
  \item Describe the relationship of this generalisation to Parikh matrices!
}

\subsubsection{Example~1: String length}

A useful example of such a morphism can express string length. We let $\Monoid =
\left(\mathbb{Z};+;0\right)$ and let $\Map(x) = 1$. Then the length of a string $s
= s_1 \RepeatSum{\Concat}  s_n$ is given by $\Map(s) = \sum \Map(s_i) = 1
\RepeatSum{+} 1$. Similarly, vectors and vector addition can be used to obtain
the regular Parikh map.

\subsubsection{Example~2: Parikh Automata}

Another application for this generalisation is to express Parikh
automata~\cite{parikh-automata}, and this is indeed what the implementation
described in Section~\ref{sec:implementation} does.

\contents{
 \item what is parikh automata
 \item Give an actual example of a parikh automata
}

\subsection{Lazy Computation of Parikh Images for Regular Languages}

We assume an NFA $\Automaton = \AutomatonTuple$ with $\NrTransitions$
transitions $\Transitions =
\Set{\EllipsisSequence{\Transitions}{\NrTransitions}}$ where we describe each
such transition $\Transitions_i$ from state $\State$ to state $\State'$ with
label $\Label \in \Alphabet$ as $\Transitions_i =
\FromLabelTo{\State}{\Label}{\State'}$. When one or more of the states and
labels of a transition are uninteresting we will omit it.

For convenience, we introduce the following supporting concepts and notations:

\begin{definition}
  A \textit{path} $\Path = \PathEnumeration$ of an automaton~$\Automaton$ with
  transitions $\Transitions$ represents a path through $\Automaton$ using
  transitions in $\Transitions$ (i.e. $\Transitions(\State_k, \Label_{k+1},
  \State_{k+1})$ holds for every $k$), passing zero or more labels $\Label_1,
  \ldots, \Label_n$. The path must begin in the initial state, i.e.~$\State_0 =
  \InitialState$. However, the ending state, $\State_n$, is not necessarily an
  accepting state.
  \end{definition}

\begin{definition}
  Moreover, we also talk about the \textit{set of paths} of an automaton,
  $\Paths(\Automaton)$, a possibly infinite (if $\Automaton$ has loops) set of
  valid paths through $\Automaton$. Additionally, we use the
  notation~$\Paths(\Automaton, \State)$ to mean all paths ending in
  state~$\State$.
\end{definition}

\begin{definition}
  The \textit{word} of a path $\WordOf(\Path) = \Label_1 \RepeatSum{\Concat} \Label_n$ is
  the word read out on its labels.
\end{definition}

\begin{definition}
  The \textit{states} of a path $\StatesOf(\Path) = \Set{\State \SuchThat
  \FromLabelTo{\State}{}{\State'} \in \Path \text{ or }
  \FromLabelTo{\State'}{}{\State} \in \Path}$ are the states visited along a
  path. Note that $\InitialState \in \StatesOf(\Path)$ for every path since all
  paths start in the initial state.
\end{definition}

\begin{definition}
  A \textit{cut}, $C$ of an automaton~$\Automaton = \AutomatonTuple$,
  $\SeparatingCut(C, \Automaton, \State)$, is a minimal set of transitions whose
  removal from~$\Transitions$ would cause~$\State \in \States$ to be unreachable
  from the initial state~$\InitialState$. $\SeparatingCut(\Automaton,
  \InitialState)$ is crucially undefined since it cannot be made unreachable
  from itself by definition.
\end{definition}

\begin{definition}
 The \textit{transition count}, $\TransitionCount(\Transition, \Path)$ is the
 number of times a transition $\Transition =
 \FromLabelTo{\State_1}{\Label}{\State_2} \in \Transitions$ appears in a path
 $p$.
\end{definition}

We then introduce the two predicates into our calculus with the following
definitions:

\begin{definition}
  The Parikh predicate, $\SinglePredicateInstance$, for some automaton
  $\Automaton = \AutomatonTuple$, modulo some map $\Map$ to a commutative monoid
  $\Monoid$ as described in Section~\ref{sec:generalised} and with a transition
  selection function $\Filter:\: \Transitions \to \Naturals$ holds when
  $\MonoidElement \in \Monoid$ is the Parikh image of $\Automaton$ modulo
  $\Map$, or more formally when there is an accepting path $\Path =
  \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ such that:
  \begin{itemize}
    \item $\Filter(\Transition) = \TransitionCount(\Transition, \Path)$
    \item $\MonoidElement = \Map(\WordOf(\Path))$
  \end{itemize}
\end{definition}

\begin{definition}
  $\Connected(\Automaton, \Filter)$ for some automaton $\Automaton =
  \AutomatonTuple$ holds when for every $\Transition =
  \FromLabelTo{\State}{}{\State'} \in \Transitions$, $\Filter(\Transition) > 0
  \implies \exists \Path \in \Paths(\Automaton)$ $\Filter(\Transition_i) > 0$
  for $\Transition_i \in p$, and $\State \in \StatesOf(p)$, or in words that
  there exists some $\Filter$-selected valid path that reaches $\Transition$'s
  starting state, $\State$. Intuitively, it represents the condition that
  $\Automaton$ is connected with respect to the selection function $\Filter$ for
  every transition. It is redundant to $\Image$ by design.
\end{definition}

Finally, we present the rules of our calculus for one automaton in Table~\ref{tbl:rules:single}. Note that we
operate on sets of symbols (terms, clauses). Additionally, we also use the
convention of splitting the formulas into linear inequalities ($\SomeInequalities$)
and any other clauses ($\SomeClause$).

We use the shorthand notation~$\Transitions_\Automaton$ to refer to the
transitions of an automaton~$\Automaton$. Additionally, for an
automaton~$\Automaton = \AutomatonTuple$ we allow mapping the selection function
like so: $\Filter(\Automaton) = \Tuple{\States, \InitialState, \AcceptingStates,
\Set{\Transition \in \Transitions \SuchThat \Filter(\Transition) > 0}}$, i.e.
$\Automaton$~with only the transitions for which~$\Filter$ is positive. In this
instance, the basis for the matched linear inequalities is
implicitly~$\SomeInequalities$.

These rules are meant to be executed in an automated theorem prover, and are
read bottom-up. Since a large part of our rules operate by adding and literally
matching linear (in)equalities in a proof goal, we use the shorthand of listing
the matched inequalities as antecedents. An example of this can be seen in
the~\Propagate{} rule.

Please note that the filtering function~$\Filter$ is evaluated symbolically in
these rules, and can in practice be read as a symbolic function from transitions
to terms (e.g.~\texttt{t} or~\texttt{t+1}), not dissimilar to~$\FlowEq$ below.
In our implementation, $\Filter$~is a vector of fresh unknown constants with the
same size as~$\Transitions$.

We assume that every rule only fires when it would add a new clause. For
example, this means that we cannot use~\Split{} to split on the same term twice.
This suggests a proof strategy where you \Propagate{} when you can, \Split{}
when you must, and \Subsume{} when neither is possible anymore.

\begin{table}[h]
\begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
  Name & Rule & Prerequisites\\
  \midrule

  % EXPAND
  \Expand & 
    \inferrule
  {\Connected(\Automaton, \Filter) \land \FlowEq(\Automaton, \Filter) \land \MonoidElement = \sum_{\Transition \in \Transitions_\Automaton} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
  {\SinglePredicateInstance, \SomeInequalities, \SomeClause} & 
  None \\

  % SPLIT
  \Split & 
  \inferrule{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0 \mid \Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) > 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
  if $\Transition \in \Transitions_\Automaton$ \\

  % PROPAGATE
  \Propagate &
  \inferrule{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause, \Filter(\Transition) = 0}{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause} &
  if $t = \FromLabelTo{\State}{}{\State'} \in \Transitions_\Automaton, \SeparatingCut(C, \Automaton, \State)$\\

  % SUBSUME
  \Subsume &
  \inferrule{\SomeInequalities,\SomeClause}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
  \Split{} and \Propagate{} cannot be applied \\
  \bottomrule
  \end{tabular}
  \caption{Derivation rules for one automaton.}\label{tbl:rules:single}
\end{table}

We use the symbolic function $\FlowEq(\Automaton, \Filter)$ that generates a set
of existentially quantified linear inequalities with the following definition:
$$
\begin{aligned}
  & \FlowEq(\Automaton, \Filter) = \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)}
\text{ $\geq 0$ if $\State \in \AcceptingStates$, $= 0$ otherwise}\\
  & \In(\State, \Filter) = \left(\text{$1$ if $\State = \InitialState$, otherwise $0$}\right) + \sum_{\Transition = \FromLabelTo{\State_0}{}{\State} \in \Transitions} \Filter(\Transition)\\
  & \Out(\State, \Filter) = \sum_{\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions} \Filter(\Transition)
\end{aligned}
$$
Where we transparently assign fresh variables to each state $\State \in \States$
and existentially quantify them in the whole clause.

Additionally, we assume the existence of a rule \PresburgerClose{},
corresponding to a sound and complete solver for Presburger formulae, modulo the
monoid~$\Monoid$.

The $\Propagate{}$ rule allows us to propagate (dis-)connectedness across
$\Automaton$. It states that we are only allowed to use transitions attached to
a reachable state, and is necessary to ensure connectedness in the presence of
cycles in~$\Automaton$.

\textsc{Expand} expands the predicate into its most basic rules; one set of
linear equations synchronising the transitions mentioned by~$\Filter$ to the
corresponding Monoid element~$\MonoidElement$, and the linear flow equations of
the standard Parikh image formulation, as described by~\FlowEq. Since
$\Connected$ and $\Image$ are partially redundant and the difference is covered
by~$\FlowEq$, we can remove the instance of~$\Image$ when applying~$\Expand$. In
this sense, we split the semantics of the $\Image$~predicate into its counting
aspect (covered by $\FlowEq$) and its connectedness aspect (covered by
$\Connected$).

Finally, $\Split{}$ allows us to branch the proof tree by trying to exclude a
contested transition from a potential solution before concluding that it must be
included. Intuitively, this is what guarantees our ability to make forward
progress by eliminating paths through~$\Automaton$.

A decision procedure for our predicate in a tableau-based automated theorem
prover would start by expanding the predicate using the $\Expand{}$~rule. A
modestly clever theorem prover would perform algebraic substitution on the
underlying constants of~$\Filter$, boiling them down to choices of branches,
which depend on one single variable, and loop transitions. This logic
corresponds to the placement of counters for optimally edge-profiling the CFG of
a program, making up a MST of the automaton~\cite{path-profiling}.

\subsection{An Example}

\subsection{Finding the Presburger representation of a Parikh Image}
\contents{
  \item lazy qe
}

\subsection{The calculus is correct}

To be as clear as possible about what it means for our calculus to be correct,
we use more exact definitions than the traditional soundness and completeness.

\subsubsection{The calculus terminates}
\subsubsection{The calculus allows no values outside the Parikh image}
\subsubsection{The calculus excludes no values in the Parikh image}

\section{Parikh Images from Products of Automata}

To generalise the calculus to calculations on products of automata, we change the main predicate to take arbitrarily many automata:
\begin{definition}
  $\ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}$
  is true exactly when the single-automaton version of the predicate would hold
  for the automaton.

  \contents{
    \item Is the enumeration correct?
    \item Do we need a new name for the predicate?
    \item Should the automata still be subscripted?
  }
\end{definition}

\begin{definition}
When computing a product of two automata, $\Automaton_1 \times \Automaton_2$, we
need to be able to reason about the sources of a transition. For that, we
introduce the relation $\ComesFrom(\Transition, \Transition',
\Automaton_1\times\Automaton_2)$. The predicate holds when $\Transition \in
\Transitions_{\Automaton_2 \times \Automaton_2}$ is created from $\Transition', \Transition''
\in \Transitions_{\Automaton_1} \cup \Transitions_{\Automaton_2}$.\Fudge{Do I need to be more specific?}
\end{definition}

  First we extend $\Expand$ to generate flow equations and instances of $\Connected$ for each automaton:
  \begin{mathpar}
    \inferrule*[left=ExpandM]
      {\Set{\Automaton_i}{\FlowEq(\Automaton_i), \Connected(\Automaton_i)}, \ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
      {\ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
  \end{mathpar}

\ExpandM{} must be applied before any other rule, like \Expand{}, but unlike \Expand{}, \ExpandM{} does not remove the $\Image$~predicate since it is needed to keep track of the product.

Then we introduce the rule $\Materialise$, used to compute a partial product between two terms:
\begin{mathpar}
  \inferrule*[left=Materialise,right=\textnormal{where $\Automaton' = \Automaton_1\times\Automaton_2$}]
    {\Set{\Filter(\Transition) = \sum \Filter(\Transition') \HoldsThat \Transition' \in \Transitions_{\Automaton'}, \ComesFrom(\Transition', \Transition, \Automaton') \SuchThat \Transition \in \Transitions_{\Automaton_1} \cup \Transitions_{\Automaton_2}},\\\ImagePredicate{\Automaton'\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
    {\ImagePredicate{\Automaton_1\times\Automaton_2\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
\end{mathpar}

Note that we implicitly map $\Filter$ to fresh terms for each transition in the
product just like before.

Finally, for instances of precisely one automaton, neither rule applies and we
perform the calculus as before.

\subsection{An Example}

\subsection{The extended calculus is also correct}
\subsubsection{The calculus terminates}
\subsubsection{The calculus allows no values outside the Parikh image}
\subsubsection{The calculus excludes no values in the Parikh image}

\section{Backjumping and Learning No-Goods}

\subsection{An Example}

\section{Implementation and Experiments}\label{sec:implementation}

We implement a calculus for Parikh automata as described in
Section~\ref{sec:parikh-automata}. The artefact submitted along with this paper
is a program that reads one or more products of one or more DFA, the
register incrementations performed along the edges of their transitions, and the
labels of the transitions as ranges of Unicode characters, along with a set of
constraints on the final values of their registers expressed as Presburger
arithmetic. We call this program \Catra{}.\footnote{If you really must read it
as an acronym, please read it as CAtegory Theory on Register Automata, or
alternatively if you find this whole paper to be too much of a theoretical
exercise as Can Anyone Think of A Real Application.}

\Catra{} is written in Scala, with the calculus described in this paper
implemented as a theory plug-in for the \Princess{} automated theorem
prover~\cite{princess}, which also performs the Presburger reasoning. For
comparison, we also provide an implementation of the baseline method
from~\cite{generate-parikh-image}, a direct translation that uses the~\Nuxmv{}
symbolic model checker~\cite{nuxmv} to solve our constraints, and \Fudge{the
approximation described in~\cite{approximate-parikh} on top of the standard
Verma back-end}. An example of an input file \Fudge{illustrating every feature
of the file format} can be seen in Listing~\ref{lst:input-example}.

\begin{lstlisting}[caption={An example input file for \Catra{}, illustrating every major syntax element. From beginning to end: synchronised (product) automata using the keyword \texttt{synchronised} (automata A and B), range labels, and their single-character shorthand syntax, register increments, and constraints on the final values of their counters.}, label=lst:input-example]
  /* Declare counters. Initially only integers, but other data-types
 could be considered as well. All counters start at zero. */
counter int x, y, z, length;

// List automata. Automata transitions are labelled with intervals,
// and can specify counter increments/decrements. A counter can only
// be used in at most one automaton.

synchronised {
    automaton A {
    init s0;

    s0 -> s1 [0, 10] { x += 1};

    accepting s1, s2;
    };

    automaton B {
    init t1;

    t1 -> t2 [11, 22] { y += 1};

    accepting t1, t2;
    };
};

automaton C {
    init s0;

    s0 -> s1 [15] { z += 1 };
    s1 -> s1 [15] { z += 1};

    accepting s1;
};

// Specify a constraint on the final values of counters; check whether
// there is a word accepted by all automata for which the final counter
// values satisfy this constraints.
constraint z = 2x + 2y && !(x = 0);
\end{lstlisting}

\contents{
  \item Do we need to say any more about the nuxmv backend?
  \item Optimisations and tricks
}

In satisfaction mode, supported by all back-ends, \Catra{} tries to satisfy the
constraints expressed by the input file, reporting satisfiable with register
assignments or unsatisfiable much like traditional~SAT- or SMT solvers would.
Additionally, \Fudge{the baseline and} our own back-end also supports generating
the Presburger formula describing the constraints of the input file \Fudge{using
the method described in Section~\ref{sec:presburger-representation}}

\subsection{\Catra{} is Versatile: Encoding Epistemic Logic Problems As Parikh Automata}

\contents{
\item Model-checking examples e.g. \cite{epistemic-logic}
}

\subsection{\Catra{} is Faster Than The Competition: Benchmarks and Their Results}

\contents{
\item length constraints
\item Parikh automata, automata with registers
\item Where do the instances come from!? Why?
\item Result table
\item Definition of winning
\item We compare to an implementation using the same automated theorem prover, the same product implementation
}

\section{Conclusions}

\contents{
  \item Future extensions: more logics
  \item Loop invariants?
  \item Integration into string solvers?
  \item We have shown fastest
  \item We have shown versatile
  \item We have shown concretely useful
}

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

%% Bibliography
\bibliography{bibliography}

%% Appendix
\appendix
\section{Appendix}

Text of appendix \ldots

\end{document}