### 2023-06-24 calculus.tex
An acceptable (but hardly great) example
This also adds propagation to 2.1, which I had forgotten.

Punching out from today after 5h 30 min.

We start by defining our calculus, \Calculus{}, for one automaton, and only
extend it to products of automata in \cref{sec:multiple}. We assume basic
familiarity with first-order logic, Presburger arithmetic, and the classical
sequent calculus. For reference, see e.g. \cite{Fitting96a}.

First, assume an NFA $\Automaton = \AutomatonTuple$ with $\NrTransitions$
transitions $\Transitions =
\Set{\EllipsisSequence{\Transition}{\NrTransitions}}$. For convenience, we then
introduce the following supporting notations:

\begin{definition}
  A \textit{path} $\Path = \PathEnumeration$ of an automaton~$\Automaton$ with
  transitions $\Transitions$ represents a path through $\Automaton$ using
  transitions in $\Transitions$ (i.e. $\Transitions(\State_k, \Label_{k+1},
  \State_{k+1})$ holds), passing zero or more labels $\Label_1, \ldots,
  \Label_n$. The path must begin in the initial state, i.e.~$\State_0 =
  \InitialState$. The end state, $\State_n$, is not necessarily accepting. If it
  is, we say the path is also \textit{accepting}.
  \end{definition}

\begin{definition}
  Moreover, we talk about the \textit{set of paths} of an automaton,
  $\Paths(\Automaton)$, a possibly infinite (if $\Automaton$ has loops) set of
  valid paths through $\Automaton$. Additionally, we use the
  notation~$\Paths(\Automaton, \State)$ to mean all paths ending in
  state~$\State$.
\end{definition}

\begin{definition}
  For a path $\Path = \Tuple{\State_0 \Label_1 \State_1 \RepeatSum{,}
  \State_n}$, its \textit{word} $\WordOf(\Path) = \Label_1 \RepeatSum{\Concat}
  \Label_n$ is the word read out on its labels.
\end{definition}

\begin{definition}
  The \textit{states} of a path $\Path = \Tuple{\State_0 \Label_1 \State_1
  \RepeatSum{,} \State_n}$, $\StatesOf(\Path) = \Set{\State_0\RepeatSum{,}
  \State_n}$ are the states visited along $\Path$. Note that $\InitialState \in
  \StatesOf(\Path)$ for every path since all paths start in the initial state.
\end{definition}

\begin{definition}
  A \textit{dominating set of transitions}, $C$ of an automaton~$\Automaton$ of
  a transition $\Transition$, written $\Dominates(C, \Automaton, \Transition)$,
  is a set of transitions from $\Automaton$ such that every accepting path $p
  \in \Automaton$ where $\Transition \in p$ contains at least one transition
  from $C$. Notably, $\Dominates(\emptyset, \Automaton, \Transition')$ for every
  dead transition $\Transition'$ and $\Dominates(Set{\Transition''}, \Automaton,
  \Transition)$.
\end{definition}

\begin{definition}
 The \textit{transition count}, $\TransitionCount(\Transition, \Path)$ is the
 number of times a transition $\Transition =
 \FromLabelTo{\State_1}{\Label}{\State_2} \in \Transitions$ appears in a path
 $p$.
\end{definition}

We then introduce the two predicates into our calculus with the following
definitions:

\begin{definition}\label{def:single-image}
  The Parikh predicate, $\SinglePredicateInstance$, for some automaton
  $\Automaton = \AutomatonTuple$, modulo some map $\Map : \Alphabet^* \to \Monoid$ to a
  commutative monoid $\Monoid$ as described in \cref{sec:generalised}
  and with a transition selection function
  $\Filter:\: \Transitions \to \Naturals$ holds when
  $\MonoidElement \in \Monoid$ is an element of the Parikh image of
  $\Language(\Automaton)$
  modulo $\Map$, or more formally when there is an accepting path
  $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ such
  that $\Filter(\Transition) = \TransitionCount(\Transition, \Path)$
  for all $t \in \Transitions$, and $\MonoidElement = \Map(\WordOf(\Path))$.
\end{definition}

\begin{definition}\label{def:connected} $\Connected(\Automaton, \Filter)$ for
  some automaton $\Automaton = \AutomatonTuple$ holds when for every
  $\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions$,
  $\Filter(\Transition) > 0 \implies \exists \Path \in \Paths(\Automaton)$
  $\Filter(\Transition_i) > 0$ for $\Transition_i \in p$, and $\State \in
  \StatesOf(p)$, or in words that there exists some $\Filter$-selected valid
  path that reaches $\Transition$'s starting state, $\State$. It represents the
  condition that $\Automaton$ is connected with respect to the selection
  function $\Filter$ for every transition. It is redundant to $\Image$ by
  design.
\end{definition}

We present the rules of \Calculus{} for one automaton in
\cref{tbl:rules:single}. The rules operate on sets of formulas, and
can be interpreted as rules of a one-sided sequent calculus, in
which all formulas are located in the antecedent~\cite{Fitting96a}.

Each rule should be read and applied bottom-up, and relates
premises~$\SomeClause_1, \ldots, \SomeClause_k$ with some
conclusion~$\SomeClause$. When constructing a proof, we start from some
root~$\SomeClause$, and then apply proof rules to the goals of the proof until a
goal can be closed, or no more rule is applicable.

A proof in which all proof goals could be closed shows that the formulas
in the root~$\SomeClause$ of the proof are inconsistent, and do not
have any solutions. A proof goal that could not be closed, but to which no
more rules are applicable, gives rise to  a solution of the formulas in the
root~$\SomeClause$. Such an open proof goal will only contain formulas
in Presburger arithmetic allowing a solution to be computed using
standard algorithms~\cite{Fitting96a}.

We use the convention of splitting the formulas into linear (in)equalities
($\SomeInequalities$) and other formulas ($\SomeClause$), and assume that
predicates $\Image$ and $\Connected$ only occur positively. Since our rules
operate by adding and matching linear (in)equalities in a proof goal, we use the
shorthand of listing the matched inequalities as antecedents e.g., in
\Propagate{}).

The filtering function~$\Filter$ is evaluated symbolically, and can in practice
be read as a function from transitions to $\Naturals$-valued terms
(e.g.~\texttt{t} or~\texttt{t+1}). In our implementation \Catra{}, described in
\cref{sec:implementation}, $\Filter$~is a vector of fresh variables with the
same size as~$\Transitions$.

We use the shorthand notation~$\Transitions_\Automaton$ to refer to the
transitions of an automaton~$\Automaton$. Additionally, for an
automaton~$\Automaton = \AutomatonTuple$ we allow mapping the selection function
like so: $\Filter(\Automaton) = \Tuple{\States, \InitialState, \AcceptingStates,
\Set{\Transition \in \Transitions \SuchThat \Filter(\Transition) > 0}}$, i.e.
$\Automaton$~with only the transitions for which~$\Filter$ is positive. In this
instance, the basis for the matched linear inequalities is
implicitly~$\SomeInequalities$. Similarly, for our commutative monoid $\Monoid =
\left(X;\MonoidOp;0_{\Monoid}\right)$ and the map into it $\Map : \Strings
\rightarrow X$, we also allow mapping over transitions:
$\Map(\FromLabelTo{\State}{\Label}{\State'}) = \Map(\Label)$.

We require that rules can only be applied when they add new formulas
on every created branch (the notion of \emph{regularity} of a proof is
required~\cite{Fitting96a}). For example, this means that \Split{} can
only be applied to proof goals that contain neither
$\Filter(\Transition) = 0$ nor $\Filter(\Transition) > 0$, and can
never be applied to split on the same term twice on the same branch.
This suggests a proof strategy where you \Propagate{} when you can, \Split{}
when you must, and \Subsume{} when neither is possible anymore.

\begin{table}
\begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
  Name & \text{Rule} & Side conditions\\
  \midrule

  % EXPAND
  \Expand & 
    \inferrule
  {\Connected(\Automaton, \Filter) ,~ \FlowEq(\Automaton, \Filter) ,~ \MonoidElement = \sum_{\Transition \in \Transitions_\Automaton} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
  {\SinglePredicateInstance, \SomeInequalities, \SomeClause} & 
  None \\[4ex]

  % SPLIT
  \Split & 
  \inferrule{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0 \mid \Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) > 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
  if $\Transition \in \Transitions_\Automaton$ \\[4ex]

  % PROPAGATE
  \Propagate & \inferrule{\Connected(\Automaton, \Filter),
  \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C},
  \SomeInequalities, \SomeClause, \Filter(\Transition) =
  0}{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat
  \Transition' \in C}, \SomeInequalities, \SomeClause} & if $\Dominates(C,
  \Automaton, \Transition) t \in \Transitions_\Automaton, $ \\[4ex]

  % SUBSUME
  \Subsume &
  \inferrule{\SomeInequalities,\SomeClause}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
  \Split{} and \Propagate{} cannot be applied \\
  \bottomrule
  \end{tabular}
  \caption{Derivation rules for one automaton.}\label{tbl:rules:single}
\end{table}
We use the symbolic function $\FlowEq(\Automaton, \Filter)$ that generates a set
of existentially quantified linear inequalities with the following definition,
where we assign fresh, existentially quantified variables to
$\FinalStateVar_\State, \Filter(\Transition)$ for every $\State \in
\AcceptingStates, \Transition \in \Transitions$:
\[
\begin{aligned}
  \FlowEq(\Automaton, \Filter) &= \sum\limits_{\State \in \AcceptingStates} \FinalStateVar_\State = 1 \land \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)} = \Sink(\State) \land
  \AndComp{\State \in \AcceptingStates}{\FinalStateVar_\State \geq 0}\\
  \Sink(\State) & = 0 \text{ if } \State \not\in \AcceptingStates, \FinalStateVar_\State \text{ otherwise.} \\
  \In(\State, \Filter) & = \StartFlow(\State) + \sum_{\Transition \in \FromLabelTo{\State'}{}{\State}} \Filter(\Transition)\\
  \StartFlow(\State)  &= 1 \text{ if } \State = \InitialState, \text{ otherwise $0$.} \\
  \Out(\State, \Filter) &= \sum_{\Transition \in \FromLabelTo{\State}{}{\State'}} \Filter(\Transition)
\end{aligned}
\]

In addition to \cref{tbl:rules:single}, we assume the existence of a rule
\PresburgerClose{}, corresponding to a sound and complete solver for Presburger
formulae, and for the elements of~$\Monoid$.

The $\Propagate{}$ rule allows us to propagate (dis-)connectedness across
$\Automaton$. It states that we are only allowed to use transitions attached to
a reachable state, and is necessary to ensure connectedness in the presence of
cycles in~$\Automaton$.

\textsc{Expand} expands the predicate into its most basic rules; one set of
linear equations synchronising the transitions mentioned by~$\Filter$ to the
corresponding Monoid element~$\MonoidElement$, and the linear flow equations of
the standard Parikh image formulation, as described by~\FlowEq. Since
$\Connected$ and $\Image$ are partially redundant and the difference is covered
by~$\FlowEq$, we can remove the instance of~$\Image$ when applying~$\Expand$. In
this sense, we split the semantics of the $\Image$~predicate into its counting
aspect (covered by $\FlowEq$) and its connectedness aspect (covered by
$\Connected$).

Finally, $\Split{}$ allows us to branch the proof tree by trying to exclude a
contested transition from a potential solution before concluding that it must be
included. Intuitively, this is what guarantees our ability to make forward
progress by eliminating paths through~$\Automaton$.

A decision procedure for our predicate in a tableau-based automated theorem
prover would start by expanding the predicate using the $\Expand{}$~rule. A
theorem prover would perform algebraic substitution on the underlying constants
of~$\Filter$, boiling them down to choices of branches, which depend on one
single variable, and loop transitions. This logic corresponds to the placement
of counters for optimally edge-profiling the CFG of a program, making up a
minimum-spanning tree of the automaton~\cite{path-profiling}.

In order to make the examples below tractable, we will assume the existence of a
rule \EquationReasoning{} that allows us to perform standard algebraic reasoning
on linear inequalities. This rule is not necessary for correctness or
completeness, but shortens the examples considerably.

\subsection{An Example}
Here we will return to $A'$ from \cref{sec:intuition} and performs the steps of
\cref{sec:intuition:algebra,sec:intuition:prop,sec:intuition:split} with the
formal calculus we just established, but exclude \cref{const:s2-in-b} and
therefore the entire automaton $B$ since we introduce support for products of
automata in \cref{sec:multiple}. Note that this example then becomes
satisfiable, as $s_1 = \Char{dd}, i = 1, s_2 = \Char{adda}, l_A = 4$ is a
satisfying assignment to
\cref{const:more-inside-than-before,const:s1-in-c-dd,const:s1-substring}.

Starting with~$A'$, $\Map$ extracts the increments of a transition (see
\cref{fig:a_1}), $h(\FromLabelTo{q_1}{c / i}{q_2}) = \Vec{v}$ for the increment
vector $\Vec{v}$, concretely $h(\FromLabelTo{I_A}{\Sigma / \left[ l_{A+}, i_+
\right]}{I_A}) = \left[ l_{A+}, i_+ \right]$. We use the same compact notation
here as in \cref{sec:intuition} to represent what is essentially a sparse vector
of $1$ and $0$ coefficients, i.e. = $\left[ l_{A+}, i_+ \right] = \left[1, 1,
0\right]$. Initially, we let $\Filter$ map to fresh variables to obtain:
\begin{equation}
    \begin{aligned}
      \Filter(\FromLabelTo{D}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A}) & = \TransitionVar_1    & 
                \Filter(\FromLabelTo{F_A}{\Sigma/\begin{bmatrix} l_{A+} \end{bmatrix}}{F_A} ) & = \TransitionVar_2 &       \FlowEq(A', \Filter) & = \begin{aligned}
                    1 + \cancel{\TransitionVar_5} &= \cancel{\TransitionVar_5} + \TransitionVar_3 + \TransitionVar_1\\
                    \TransitionVar_4 &= \TransitionVar_1 \\
                    \TransitionVar_3 + \TransitionVar_1 + \cancel{\TransitionVar_2} &\geq \cancel{\TransitionVar_2}  \\
                  \end{aligned} \\
      \Filter(\FromLabelTo{I_A}{c/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A} ) & =  \TransitionVar_3 & 
                \Filter(\FromLabelTo{I_A}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{D}) & = \TransitionVar_4 &\\
      \Filter(\FromLabelTo{I_A}{\Sigma/\begin{bmatrix} l_{A+} i_+\end{bmatrix}}{I_A}) & = \TransitionVar_5 & 
    \end{aligned}
\end{equation}\label{eq:single-filter}

Like in \cref{sec:a_1,sec:intuition:algebra}, we apply reasoning by equivalence
  to end up with fixed values for several of the transition variables
  $\TransitionVar_k$, and eventually we use \cref{const:more-inside-than-before}
  and \cref{const:something-before-and-after} to conclude that the transition
  directly from state $I_{A}$ to $F_{A}$ is incompatible with the constraint.

  Then, following \cref{sec:intuition:subsume}, we remove the $\Connected{}$
   predicate using \Subsume{} since no transition can can become disconnected
   (and in fact never could for $A'$, since all looping transitions are attached
   to the initial and single accepting state and therefore always must be
   reachable from any accepting path). After this, the rest of the reasoning can
   be continued without any part of our calculus, since all values of the
   existentially quantified transition variables
   $\TransitionVar_1,\ldots,\TransitionVar_5$ compatible with
   \cref{const:more-inside-than-before,const:s1-in-c-dd,const:s1-substring,const:something-before-and-after}
   are represented by the flow equations introduced by $\Expand{}$. The
   derivation is shown in \cref{fig:derivation:single}.

\begin{figure}
  \centering
\begin{prooftree}
  \hypo[]{{\text{Regular solving finishes the proof}}}
  \infer1[]{
    \begin{aligned}
    \TransitionVar_1 = 1 \land 
    i = \TransitionVar_5 \land 
    n = 2 \land \\
    2 \geq i \land l_A - i - 2 > 0 
    \land \TransitionVar_4 = 1 
    \land \TransitionVar_2 = l_A - i - 2 
    \land \TransitionVar_3 = 0 \\
    \end{aligned}}
  \infer1[\Subsume: \cref{sec:intuition:subsume}]{
    \begin{aligned}
    \Connected(A', \Filter) \land 
    \TransitionVar_1 = 1 \land 
    i = \TransitionVar_5 \land 
    n = 2 \land 
    2 \geq i \land l_A - i - 2 > 0  \\
    \land \TransitionVar_4 = 1 
    \land \TransitionVar_2 = l_A - i - 2 
    \land \TransitionVar_3 = 0 \\
    \end{aligned}
  }
  \infer1[\EquationReasoning{}: $\TransitionVar_1 = 1$ \cref{sec:a_1}]{
    \begin{aligned}
    \Connected(A', \Filter) \land \\
    i = \TransitionVar_5 \land 
    n = 2\TransitionVar_1 \land n \geq i \land l_A - i - n > 0 
    \land \TransitionVar_4 = \TransitionVar_1 \\
    \land \TransitionVar_2 = l_A - i - n 
    \land \TransitionVar_3 = 1 - \TransitionVar_1 
    \end{aligned}
  }
  \infer1[\EquationReasoning{}: $=$-elimination]{
    \begin{aligned}
    \Connected(A', \Filter) \land \\
    i = \TransitionVar_5 \land 
    n = \TransitionVar_4 + \TransitionVar_1 \land n \geq i \land l_A - i - n > 0 \\
    \land \TransitionVar_4 = \TransitionVar_1 
    \land \TransitionVar_2 = l_A - i - n 
    \land \TransitionVar_3 = 1 - \TransitionVar_1 
    \end{aligned}
  }
  \infer1[\EquationReasoning{}: expand sum]{
    \begin{aligned}
        i = \TransitionVar_5, l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k \land n = \TransitionVar_4 + \TransitionVar_1 \land n \geq i \land l_A - i - n > 0
    \land \Connected(A', \Filter)
    \end{aligned}
  }
  \infer1[\EquationReasoning{}: expand sum]{
    \Connected(A', \Filter) \land
    \left[l_A, i, n\right] =
    \sum\limits_{\Transition \in \Transitions_{A'} } \Map(\Transition) \cdot \Filter(\Transition)
  }
  \infer1[\Expand{}]{\Image{}_{A', \Map}(\Filter, \left[l_A, i, n\right]) \land n \geq i \land l_A - i - n > 0}
\end{prooftree}
\caption{A derivation tree for an application of
\cref{const:s1-in-c-dd,const:s1-substring,const:something-before-and-after,const:more-inside-than-before}
from \cref{ex:string-constraints}, corresponding to handling the Parikh
automaton $A'$ of \cref{fig:aut_a}. Note that the flow equations of
\cref{eq:single-filter} are left out of the tree in order to keep it
manageable.}\label{fig:derivation:single}
\end{figure}

\subsection{Correctness of \Calculus{}}\label{sec:single:correct}

Our correctness proof of \Calculus{} consists of two main parts: we
first show that the construction of a proof always terminates, and
then that each of the proof rules in \cref{tbl:rules:single} is
an equivalence transformation, i.e., does not change the set of
satisfying assignments of a formula. In combination, those two results
immediately imply that \Calculus{} gives rise to a decision procedure.

\subsubsection{\Calculus{} terminates}
\begin{lemma}\label{lma:single-terminates}
  Suppose $\SomeClause{}$ is a set of formulas in which the predicates
  $\Image$ and $\Connected$ only occur positively. There is no
  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
  $P_i$ by applying one of the rules in \cref{tbl:rules:single}.
\end{lemma}

\begin{proof}
  The rule~\Expand{} can only be applied finitely often, since each
  application removes one $\Image$ predicate, and none of the rules
  introduce new instances of the predicate. The rule~\Subsume{} can
  only be applied finitely often, since it strictly decreases the
  combined number of $\Image$ and $\Connected$ predicates in sets of
  formulas, and none of the rules increases that number.

  To show termination of \Split{} and \Propagate{}, observe that the
  $\Filter$ in a predicate~$\Connected(\Automaton, \Filter)$ is never
  updated on a proof branch, which means that the set of terms
  $\Filter(t)$ for $t \in \Automaton$ on every branch is finite. Each
  application of \Split{} and \Propagate{} adds a new
  formula~$\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$
  to a proof goal, which can only happen finitely often.
\end{proof}

\subsubsection{The rules in \cref{tbl:rules:single} are solution-preserving}

\begin{lemma}\label{lma:single-correct}
  Consider an application of one of the rules in
  \cref{tbl:rules:single}, with
  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
  conclusion~$\SomeClause$. An assignment~$\beta$ satisfies the
  conclusion~$\SomeClause$ if and only if it satisfies one of the
  premises~$\SomeClause_i$.
\end{lemma}

\begin{proof}
  This property has to be shown by analysing the possible applications
  of each proof rule.

  \Expand{} unfolds the definition of the $\Image$ predicate. To show
  that the rule is solution-preserving we prove the equivalence of the
  upper and lower sets of formulas:
  \begin{itemize}
  \item Assume that $\beta$ satisfies the conclusion, which means that
    there is some accepting path
    $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ with
    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
    \Path)$ and $\val_\beta(\MonoidElement) =
    \Map(\WordOf(\Path))$. Since immediately implies that $\beta$
    satisfies $\Connected(\Automaton, \Filter)$, since a path is
    connected, and $\FlowEq(\Automaton, \Filter)$ since an accepting
    path satisfies the flow equations. The
    equation~$\MonoidElement = \sum_{\Transition \in
      \Transitions_\Automaton}\Filter(\Transition) \cdot
    \Map(\Transition)$ holds because of
    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
    \Path)$.
  \item Assume that $\beta$ satisfies the premise, which implies that
    $\val_\beta(\Filter)$ describes a consistent, connected flow of
    the automaton. By the same argument as in
    \cite{generate-parikh-image}, this flow
    can be mapped to an accepting path~$\Path$ of $\Automaton$ such
    that each transition~$\Transition$ occurs on $\Path$ exactly
    $\val_\beta(\Filter(\Transition))$ times. Together with the equation
    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
    \Path)$, this implies that $\beta$ satisfies $\SinglePredicateInstance$.
  \end{itemize}

  In \Split{}, we make use of the fact that $\Filter(\Transition)$ is
  $\Naturals$-valued by definition. For any $\beta$, clearly exactly
  one of $\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$ will
  be satisfied, implying the property.

  For \Propagate{}, suppose that $\Dominates(C, \Automaton, \Transition)$, which
  means that every accepting path containing~$\Transition$ contains at least one
  of the transitions in $C$. For a $\beta$ satisfying $\Connected(\Automaton,
  \Filter)$. This means that every accepting path $p$ where $\Transition \in p$
  has at least one transition $\Transition' \in C$ such that
  $\Filter(\Transition') = 0$. This means that also
  $\val_\beta(\Filter(\Transition)) = 0$ has to hold since no unbroken path
  containing $\Transition$ exists.

  Finally, for \Subsume{}, observe that if \Split{} cannot be applied,
  then a goal must contain~$\Filter(\Transition) = 0$ or
  $\Filter(\Transition) > 0$ for every $\Transition$. In case the
  formulas in $\SomeInequalities$ are inconsistent, an application of
  \Subsume{} is trivially solution-preserving; therefore assume that
  $\SomeInequalities$ is consistent, which means that it contains
  exactly one of $\Filter(\Transition) = 0$ or
  $\Filter(\Transition) > 0$ for each $\Transition$. Since
  \Propagate{} is not applicable, the transitions~$\Transition$ with
  $\Filter(\Transition) > 0$ must form a connect sub-graph of the
  automaton; this means that $\Connected(\Automaton, \Filter)$ is
  redundant as it is implied by $\SomeInequalities$.
\end{proof}