### 2023-07-11 preliminaries.tex
some clean-up

\subsection{Monoids}

% AS: reviewer comment #54 complains that this is basic, are they right?

% PR: sure, this is rather basic; but I think we should still include
% those definitions

A monoid $\Monoid = \Tuple{X;\MonoidOp;0_{\Monoid}}$ is an algebraic structure
consisting of the non-empty carrier set of elements, $X$; an associative binary operation
$X \times X \rightarrow X$ denoted as $\MonoidOp$, that is where for all $a, b,
c \in X$, $(a \MonoidOp b) \MonoidOp c = a \MonoidOp (b \MonoidOp c)$; 
and an identity element $0_{\Monoid} \in X$ such that
$0_{\Monoid} \MonoidOp a = a \MonoidOp 0_{\Monoid} =   a$ for every $a \in X$.
We sometimes use integer multiplication to represent a repeated application of
$\MonoidOp$, e.g. $3a = a \MonoidOp a \MonoidOp a$, for $a \in X$. $\Monoid$ is
called \textit{commutative} if $\MonoidOp$ also commutes, that is if $a
\MonoidOp b = b \MonoidOp a$ for all $a, b \in X$.

A \textit{homomorphism} is a structure-preserving map between two
monoids $\Monoid = \Tuple{X;\MonoidOp;0_{\Monoid}}$ and
$\Monoid' = \Tuple{X';\MonoidOp';0_{\Monoid'}}$, that is a map
$\Map : X_1 \rightarrow X_2$ such that
$\Map(a \MonoidOp b) = \Map(a) \MonoidOp' \Map(b)$ and
$\Map(0_{\Monoid}) = 0_{\Monoid'}$.

\subsection{Languages, Finite-state Automata and their Products}

We define an alphabet as a finite set of symbols $\Alphabet$ with words $\Strings$, and
the concatenation operation as $s_1 \Concat{} s_2$ over two strings $s_1, s_2$.
Note that $\Strings = \Tuple{\Alphabet;\Concat{};\epsilon}$, is a
non-commutative monoid, referred to as the free monoid on $\Alphabet$.
The string length
function, $\Length{s}$ is an example of a homomorphism between $\Strings$
and~$\mathbb{Z}$.

A \emph{finite-state automaton}~$\Automaton$ with alphabet~$\Alphabet$ is a
tuple~$\AutomatonTuple$, where $\States$ is the set of states,
$\InitialState$ the initial state, $\AcceptingStates$ the set of
accepting states, and
$\Transitions \subseteq \States \times \Alphabet \times \States$ the
transition relation.  We write a transition
$\Transition = \Tuple{\State, \Label, \State'} \in \Transitions$ as
$\Transition = \FromLabelTo{\State}{\Label}{\State'}$.  Similarly, we
use the notation $\FromLabelTo{\State}{}{}$ to refer to the set of
transitions starting in $\State$, and $\FromLabelTo{}{}{\State}$ to
refer to the set of transitions coming into $\State$, whenever the
automaton is clear from the context. An accepting run of $\Automaton$
is a sequence
$\InitialState = q_0 \xrightarrow{l_1} q_1 \xrightarrow{l_2} \cdots
\xrightarrow{l_k} q_k \in \AcceptingStates$, and the accepted word is
$l_1 l_2 \ldots l_k \in \Strings$. The set of words accepted by an
automaton~$\Automaton$ is denoted by
$\Language(\Automaton) \subseteq \Strings$.

We will let variables $\Transition, \Transition', \Transition_1, \ldots,
\Transition_n$, etc., denote transitions, $\State, \ldots, \State_n$ states, and
$\Automaton, \ldots, \Automaton_n$ automata, and use subscript indexing
($\Transitions_\Automaton$) to refer to the transitions, states, etc., of a given
automaton.

The \emph{product} of two
automata~$\Automaton_1 = \Tuple{\States^1, \InitialState^1,
  \AcceptingStates^1, \Transitions^1}$ and
$\Automaton_2 = \Tuple{\States^2, \InitialState^2,
  \AcceptingStates^2, \Transitions^2}$ is the automaton
\begin{align*}
\Automaton_1 \times \Automaton_2 &=
\Tuple{\States_1 \times \States_2, \STuple{\InitialState^1, \InitialState^2},
\AcceptingStates^1 \times \AcceptingStates^2,
  \Transitions}\\
  \text{with}\quad \Transitions &= \{\STuple{\STuple{q,q''}, l, \STuple{q',q'''}} \mid \STuple{q, l, q'} \in \Transitions^1, \STuple{q'', l, q'''} \in \Transitions^2\}~.
\end{align*}
The product automaton runs $\Automaton_1$ and $\Automaton_2$ in
parallel on an input and only accepts the input if both automata would
do so; we have $\Language(\Automaton_1 \times \Automaton_2) =
\Language(\Automaton_1) \cap \Language(\Automaton_2)$.

\subsection{The Parikh Map and its Image}
Formally, the \textit{Parikh map} over an alphabet $\Alphabet=
\left\{a_1, \ldots, a_k \right\}$ is defined as in \cite{kozen}:
$$
\begin{aligned}
& \ParikhMap: \MapFromTo{\Strings}{\natural^k} \\
& \ParikhMap(s) = \VectorLiteral{\#a_1(s), \#a_2(s), \ldots, \#a_k(s)}
\end{aligned}
$$

That is, $\ParikhMap(s)$ is a vector of the number of occurrences of each
character in the language for a given string $s$. For example, for  $\Alphabet =
\Set{a, b}$, we would have $\ParikhMap(abb) = \VectorLiteral{1, 2}$.

We define the image of this map, the \textit{Parikh image}, of some subset of
the language $\Language \subseteq \Strings$ as:
\[
\ParikhMap(\Language) = \Set{\ParikhMap(x) \SuchThat x \in \Language}
\]

Thus, we would have $\ParikhMap(\left\{ab, abb\right\}) = \left\{\left[1,
1\right], \left[1, 2\right]\right\}$. We also sometimes use the standard
notation $\#l(w)$ to talk about an individual letter $l$ in a word $w$. For
example, for the Parikh vector above, we would have $\CountOf{a} = 1$.

% AS: Reviewers suggest we drop this:
% PR, shortened a bit
Parikh's theorem states that any context-free language has a Parikh-equivalent
regular language (c.f.~\cite{construction} for a construction of such automata
from context-free grammars and~\cite{bounds} for bounds on its size).
The Parikh image is therefore a semi-linear set and Presburger-definable.
%However,
%there are languages that are not context-free that also have semilinear images
%under~$\ParikhMap$ (e.g. $\ParikhMap(\Set{a^nb^nc^n \SuchThat n \geq 0}) =
%\ParikhMap((abc)^*) = \CountOf{a} = \CountOf{b} = \CountOf{c} \land \CountOf{a}
%\geq 0$). This means they can be represented as a quantifier-free Presburger
%formula.
%
While Parikh's theorem applies to arbitrary context-free languages, in this
paper we focus only on regular languages.

\subsection{The Parikh image of a regular language expressed in Presburger arithmetic}
\label{sec:verma}

Since Parikh images are semilinear, any Parikh image can be written as a set of
linear equations. The version for context-free grammars presented in
\cite{generate-parikh-image} can be straightforwardly adapted for use with a
product of regular languages or Parikh automta. For an intuition, the approach
consists of first computing the product, then assigning each state and
transition an existentially quantified non-negative integer variable, then
describing all paths through the automaton through two sets of equations: one
that preserves the incoming flow across the variables of a state's incoming
transitions is equal to the flow across its outgoing transitions, and one set of
constraints that enforces coherence in the presence of loops by ordering states
by distance in a spanning tree rooted in the initial state.

We refer to this model as the baseline approach, though we also apply
optimisations as described in \cref{sec:implementing-baseline}. The calculus
introduced in this paper, by contrast, lazily enforces the connectedness
constraint while also interleaving the computation of products of automata and
propagating information between the steps to reduce the amount of work that
needs to be done.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: