### 2023-07-11 calculus.tex
restructuring

We start by defining our calculus, \Calculus{}, for one automaton, and only
extend it to products of automata in \cref{sec:multiple}.
%
Assume an automaton $\Automaton = \AutomatonTuple$ with $\NrTransitions$
transitions $\Transitions =
\Set{\EllipsisSequence{\Transition}{\NrTransitions}}$. We use
the notations introduced in \cref{sec:languages}. For convenience, we
introduce the following additional notions:

\begin{definition}
  The \textit{transition count},
  $\TransitionCount(\Transition, \Path)$ is the number of times a
  transition
  $\Transition = \FromLabelTo{\State}{\Label}{\State'} \in
  \Transitions$ appears on a path $p$. A \emph{transition selection
    function} is a function~$\Filter:\: \Transitions \to \Naturals$
  labelling every transition~$\Transition \in \Transitions$ with a
  non-negative number.
\end{definition}

We introduce the two predicates that will be used by our calculus,
with the following definitions:
%
\begin{definition}\label{def:single-image}
  The Parikh predicate, $\SinglePredicateInstance$ holds for some
  automaton $\Automaton = \AutomatonTuple$, some map
  $\Map : \Alphabet^* \to \Monoid$ to a commutative monoid $\Monoid$,
  some transition selection function
  $\Filter:\: \Transitions \to \Naturals$, and some monoid
  element~$\MonoidElement \in \Monoid$ if $\MonoidElement$ is an
  element of the Parikh image of $\Language(\Automaton)$ modulo
  $\Map$, or more formally, when there is an accepting path
  $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ such
  that $\Filter(\Transition) = \TransitionCount(\Transition, \Path)$
  for all $t \in \Transitions$, and
  $\MonoidElement = \Map(\WordOf(\Path))$.
\end{definition}

\begin{definition}\label{def:connected} $\Connected(\Automaton, \Filter)$
  holds for some automaton $\Automaton = \AutomatonTuple$ and
  transition selection function
  $\Filter:\: \Transitions \to \Naturals$
  if for every
  $\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions$,
  if $\Filter(\Transition) > 0$ there is some $\Filter$-selected
  accepting path that visits $\Transition$'s starting state~$\State$. More formally,
  if $\Filter(\Transition) > 0$ then there is some
  path~$\Path \in \Paths(\Automaton)$ with
  $\Filter(\Transition') > 0$ for every $\Transition' \in p$, and $\State \in
  \StatesOf(p)$ such that $p$ ends in an accepting state $\State' \in \AcceptingStates$. The predicate~$\Connected(\Automaton, \Filter)$ represents the
  condition that $\Automaton$ is connected under the selection
  function $\Filter$ for every transition, and is implied by
  $\SinglePredicateInstance$.
\end{definition}

\begin{table}
\begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
  Name & \text{Rule} & Side conditions\\
  \midrule

  % EXPAND
  \Expand & 
    \inferrule
  {\Connected(\Automaton, \Filter) ,~ \FlowEq(\Automaton, \Filter) ,~ \MonoidElement = \sum_{\Transition \in \Transitions_\Automaton} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
  {\SinglePredicateInstance, \SomeInequalities, \SomeClause} & 
  None \\[4ex]

  % SPLIT
  \Split & 
  \inferrule{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0 \mid \Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) > 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
  if $\Transition \in \Transitions_\Automaton$ \\[4ex]

  % PROPAGATE
  \Propagate & \inferrule{\Connected(\Automaton, \Filter),
  \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C},
  \SomeInequalities, \SomeClause, \Filter(\Transition) =
  0}{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat
  \Transition' \in C}, \SomeInequalities, \SomeClause} & if
                                                         $t \in \Transitions_\Automaton$ and\newline $\Dominates(C,
  \Automaton, \Transition)$ \\[4ex]

  % SUBSUME
  \Subsume &
  \inferrule{\SomeInequalities,\SomeClause}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
  \Split{} and \Propagate{}\newline cannot be applied \\
  \bottomrule
  \end{tabular}
  \caption{Derivation rules for one automaton.}\label{tbl:rules:single}
\end{table}

The rules of \Calculus{} for one automaton are given in
\cref{tbl:rules:single}. The rules operate on sets of formulas and can
be interpreted as rules of a one-sided sequent calculus, in which all
formulas are located in the antecedent~\cite{Fitting96a}.
%
The rules relate premises~$\SomeClause_1, \ldots, \SomeClause_k$ with
some conclusion~$\SomeClause$. 
%
When constructing a proof, we start with some root~$\SomeClause$, and
then apply proof rules to the goals of the proof, in bottom-up
direction, until all goals are closed, or no more rule is applicable.

A proof in which all proof goals are closed shows that the formulas
in the root~$\SomeClause$ of the proof are inconsistent, and do not
have any solutions. A proof goal that cannot be closed, but to which no
more rules are applicable, gives rise to  a solution of the formulas in the
root~$\SomeClause$. Such an open proof goal will only contain formulas
in Presburger arithmetic allowing a solution to be computed using
standard algorithms~\cite{DBLP:books/daglib/0022394}.

We use the convention of splitting the formulas in proof goals into
linear (in)equalities ($\SomeInequalities$) and other formulas
($\SomeClause$), and assume that predicates $\Image$ and $\Connected$
only occur positively.
%Since our rules operate by adding and matching
%linear (in)equalities in a proof goal; we use the shorthand of listing
%the matched inequalities as antecedents, e.g., in \Propagate{}.
%
The transition selection function~$\Filter$ is represented
symbolically and can, in practice, be read as a function from
transitions to $\Naturals$-valued terms (e.g.~\texttt{t}
or~\texttt{t+1}). In our implementation \Catra{}, described in
\cref{sec:implementation}, $\Filter$~is a vector of fresh variables
with the same size as~$\Transitions$.

To ensure termination, rules can only be applied when they add new formulas
on every created branch (the notion of \emph{regularity} of a proof is
required~\cite{Fitting96a}). For example, this means that \Split{} can
only be applied to proof goals that contain neither
$\Filter(\Transition) = 0$ nor $\Filter(\Transition) > 0$, and can
never be applied to split on the same term twice on the same branch.

\medskip
%
The rule~\textsc{Expand} expands an $\SinglePredicateInstance$
predicate into the more basic predicate
$\Connected(\Automaton, \Filter)$, as well as linear equations
relating the transitions mentioned by~$\Filter$ with the monoid
element~$\MonoidElement$, and linear flow equations described
by~\FlowEq{} (below). Since $\Connected$ and $\Image$ are partially
redundant and the difference is covered by~$\FlowEq$, we can remove
the instance of~$\Image$ when applying~$\Expand$. In this sense, we
split the semantics of the $\Image$~predicate into its counting aspect
(covered by $\FlowEq$) and its connectedness aspect (covered by
$\Connected$).

In \textsc{Expand}, we use the shorthand
notation~$\Map(\FromLabelTo{\State}{\Label}{\State'}) = \Map(\Label)$,
i.e., we allow the homomorphism~$h$ to be applied also to
transitions~$\Transition$.
%
The predicate $\FlowEq(\Automaton, \Filter)$ represents the flow
equations to be generated when expanding
$\SinglePredicateInstance$. We assume that each application of the
predicate introduces fresh integer variables~$f_q$ for every accepting
state~$q \in \AcceptingStates$, and define:
%
\[
\begin{aligned}
  \FlowEq(\Automaton, \Filter) &~=~
   \AndComp{\State \in \AcceptingStates}{\FinalStateVar_\State \geq 0}
   \land
   \sum\limits_{\State \in \AcceptingStates} \FinalStateVar_\State = 1 \land \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)} = \Sink(\State)\\
  \Sink(\State) & ~=~ 0 \text{ if } \State \not\in \AcceptingStates, \FinalStateVar_\State \text{ otherwise.} \\
  \In(\State, \Filter) & ~=~ \StartFlow(\State) + \sum_{\Transition \in\, (\FromLabelTo{}{}{\State})} \Filter(\Transition)\\
  \StartFlow(\State)  &~=~ 1 \text{ if } \State = \InitialState, \text{ otherwise $0$.} \\
  \Out(\State, \Filter) &~=~ \sum_{\Transition \in\, (\FromLabelTo{\State\,}{}{})} \Filter(\Transition)
\end{aligned}
\]

The rule~$\Split{}$ allows us to branch the proof tree by trying to exclude a
contested transition from a potential solution before concluding that it must be
included. Intuitively, this is what guarantees our ability to make forward
progress by eliminating paths through~$\Automaton$.

The $\Propagate{}$ rule allows us to propagate (dis-)connectedness
across $\Automaton$. It states that we are only allowed to use
transitions attached to a reachable state, and is necessary to ensure
connectedness in the presence of cycles in~$\Automaton$. The rule
makes use of the notion of \emph{dominating} sets of transitions, defined
as follows:
%
\begin{definition}
  A set~$C$ of transitions of an automaton~$\Automaton$ \emph{dominates}
  a  transition~$\Transition$, written
  $\Dominates(C, \Automaton, \Transition)$, if
  every accepting path $p$ of $\Automaton$
  with $\Transition \in p$ contains at least one transition from
  $C$. Notably, $\Dominates(\emptyset, \Automaton, \Transition)$ for
  every unreachable transition $\Transition$, and
  $\Dominates(\Set{\Transition}, \Automaton, \Transition)$ for every
  transition~$\Transition$.
\end{definition}

Finally, the rule~\Subsume{} can be applied when the connectedness of
an automaton has been ensured by exhaustive application of the other
rules. This suggests a proof strategy where you \Propagate{} when you
can, \Split{} when you must, and \Subsume{} when neither is possible
anymore.

In addition to \cref{tbl:rules:single}, we assume the existence of a rule
\PresburgerClose{}, corresponding to a sound and complete solver for Presburger
formulae, and for constraints over~$\Monoid$.

\medskip
%
A decision procedure would start from one or multiple
predicates~$\SinglePredicateInstance$ to be satisfied, possibly in
combination with other constraints about $m$. It would then first
expand the predicates using the $\Expand{}$~rule, and subsequently
apply the other rules to search for a solution.

As illustrated in \cref{sec:intuition}, a decision procedure can
also perform arithmetic rewriting of the occurring terms and
equations. Such reasoning is not necessary for correctness or
completeness, but it shortens the examples considerably; we
will therefore assume the existence of a
rule \EquationReasoning{} that allows us to perform standard algebraic reasoning
on linear arithmetic constraints.

%, boiling them down to choices of branches, which depend
%on one single variable and loop transitions. This logic corresponds to
%the placement of counters for optimally edge-profiling the CFG of a
%program, making up a minimum-spanning tree of the
%automaton~\cite{path-profiling}.

\subsection{An Example}\label{sec:single:example}
Here we will return to $\mathcal{A}'$ from \cref{sec:intuition} and perform the steps of
\cref{sec:intuition:algebra,sec:intuition:split} with the
formal calculus we just established, but exclude \cref{const:s2-in-b} and
therefore the entire automaton $\mathcal{B}$, since we introduce support for products of
automata in \cref{sec:multiple}. The example then becomes
satisfiable, as $s_1 = \Char{dd}, i = 1, s_2 = \Char{adda}, l_A = 4$ is a
satisfying assignment to
\cref{const:more-inside-than-before,const:s1-in-c-dd,const:s1-substring}.

Starting with~$\mathcal{A}'$, $\Map$ extracts the increments of a transition, $h(\FromLabelTo{q_1}{c / i}{q_2}) = \Vec{v}$ for the increment
vector $\Vec{v}$, concretely $h(\FromLabelTo{I_A}{\Sigma / \left[ l_{A+}, i_+
\right]}{I_A}) = \left[ l_{A+}, i_+ \right]$. We use the same compact notation
here as in \cref{sec:intuition} to represent what is essentially a sparse vector
of $1$ and $0$ coefficients, i.e. $\left[ l_{A+}, i_+ \right] = \left[1, 1,
0\right]$. Initially, we let $\Filter$ map to fresh variables to obtain, after some simplifications:
\begin{equation}\label{eq:single-filter}
  \begin{array}{rl@{\qquad}rl@{\qquad}rl}
    \Filter(\FromLabelTo{D}{d/[l_{A+}, n_+]}{F_A})
    &= \TransitionVar_1
    &
    \Filter(\FromLabelTo{F_A}{\Sigma/[ l_{A+} ]}{F_A} )
    &= \TransitionVar_2   
    &
    \Filter(\FromLabelTo{I_A}{c/[l_{A+}, n_+]}{F_A} ) 
    &=  \TransitionVar_3 
    \\
    \Filter(\FromLabelTo{I_A}{d/[l_{A+}, n_+]}{D}) 
    &= \TransitionVar_4
    &
    \Filter(\FromLabelTo{I_A}{\Sigma/[ l_{A+}, i_+]}{I_A}) 
    &= \TransitionVar_5  \\[3ex]
%
  \multicolumn{6}{c}{
    \FlowEq(\mathcal{A}', \Filter)  = \begin{aligned}
                       1 + \cancel{\TransitionVar_5} &= \cancel{\TransitionVar_5} + \TransitionVar_4 + \TransitionVar_3\\
                    \TransitionVar_4 &= \TransitionVar_1 \\
                    \TransitionVar_3 + \TransitionVar_1 + \cancel{\TransitionVar_2} &= \cancel{\TransitionVar_2} + 1  \\
                    \end{aligned}}
  \end{array}
\end{equation}

The proof tree is shown in \cref{fig:derivation:single}.  Like in
\cref{sec:a_1,sec:intuition:algebra}, by arithmetic reasoning the
calculus ends up with fixed values for several of the transition
variables $\TransitionVar_k$, and eventually concludes that the
transition directly from state $I_{A}$ to $F_{A}$ (variable~$x_3$)
is incompatible with
the constraint.  It is then possible to remove the $\Connected{}$
predicate using \Subsume{}, since \Propagate{} is not able to infer
further constraints, and no non-trivial applications of \Split{}
remain to be done.  After this, by further arithmetic reasoning we can
derive a solution~$l_A = 4, i = 1, n = 2$, and conclude that the root
constraint is satisfiable.  To obtain values for the string
variables~$s_1, s_2$ corresponding to the solution, one can
construct an accepting path of the automaton with each transition~$t$
taken $\Filter(t)$ times.

\iffalse

the rest of the reasoning can be
continued without any part of \Calculus{} since all values of the
existentially quantified transition variables
$\TransitionVar_1,\ldots,\TransitionVar_5$ compatible with
\cref{const:more-inside-than-before,const:s1-in-c-dd,const:s1-substring,const:something-before-and-after}
are represented by the flow equations introduced by $\Expand{}$.
\fi

\begin{figure}[tb]
  \centering
\begin{prooftree}
  \hypo[]{$\vdots$}
  \infer1[]{
      x_1 = 1 \land
      x_2 = l_A - 3 \land
      x_3 = 0 \land 
      x_4 = 1 \land
      x_5 = 1 \land
      n = 2 \land i = 1
    \land l_A > 3
  }
  \infer1[\Subsume]{
    \begin{array}{c}
      \Connected(A', \Filter) \land \mbox{}\\
      x_1 = 1 \land
      x_2 = l_A - 3 \land
      x_3 = 0 \land 
      x_4 = 1 \land
      x_5 = 1 \land
      n = 2 \land i = 1
    \land l_A > 3
    \end{array}
%    \begin{aligned}
%    i = \TransitionVar_5 \land 
%    n = 2 \land \\
%    0 < i < 2
%    \land l_A - i - 2 > 0 
%    \land \TransitionVar_2 = l_A - i - 2 
%    \land \TransitionVar_3 = 0 \\
%  \end{aligned}
  }
  \infer1[\EquationReasoning: $=$-simplification]{
    \begin{array}{c}
      \Connected(A', \Filter) \land\mbox{}\\
      x_1 = n - 1 \land
      x_2 = l_A - n - i \land
      x_3 = 2 - n \land 
      x_4 = n - 1 \land
      x_5 = i \land \mbox{}\\
      n = 2 \land i = 1
    \land l_A > 3
    \end{array}
%    \begin{aligned}
%    \Connected(\mathcal{A}', \Filter) \land 
%    \TransitionVar_1 = 1 \land 
%    i = \TransitionVar_5 \land 
%    n = 2 \land \\
%    0 < i < 2 
%    \land l_A - i - 2 > 0
%    \land \TransitionVar_2 = l_A - i - 2 
%    \land \TransitionVar_3 = 0
%    \end{aligned}
  }
  \infer1[\EquationReasoning{}: $>$-reasoning]{
    \begin{array}{c}
      \Connected(A', \Filter) \land\mbox{}\\
      x_1 = n - 1 \land
      x_2 = l_A - n - i \land
      x_3 = 2 - n \land 
      x_4 = n - 1 \land
      x_5 = i \land \mbox{}\\
      n > i 
    \land l_A - i - n > 0
    \land i > 0
    \end{array}
%    \begin{aligned}
%    \Connected(\mathcal{A}', \Filter) \land
%    i = \TransitionVar_5 \land 
%    n = 2\TransitionVar_1 
%    \land n > i > 0 \\
%    \land l_A - i - n > 0 
%    \land \TransitionVar_2 = l_A - i - n 
%    \land \TransitionVar_3 = 1 - \TransitionVar_1 
%    \end{aligned}
  }
  \infer1[\EquationReasoning{}: $=$-simplification]{
    \begin{array}{c}
      \Connected(A', \Filter) \land\mbox{}\\
      1 = x_4 + x_3 \land x_4 = x_1 \land x_3 + x_1 = 1 \land \mbox{}\\
      l_A = x_1 + x_2 + x_3 + x_4 + x_5 \land
      i = x_5 \land 
      n = x_1 + x_3 + x_4 \land \mbox{}\\
      n > i 
    \land l_A - i - n > 0
    \land i > 0
    \end{array}
%    \begin{aligned}
%    \Connected(\mathcal{A}', \Filter)  
%    \land i = \TransitionVar_5 
%    \land n = \TransitionVar_4 + \TransitionVar_1 
%    \land n > i > 0
%    \land l_A - i - n > 0
%    \land \TransitionVar_4 = \TransitionVar_1 \\
%    \land \TransitionVar_2 = l_A - i - n 
%    \land \TransitionVar_3 = 1 - \TransitionVar_1
%    \land 1 = \TransitionVar_3 + \TransitionVar_1 
%    \land i > 0
%    \land \TransitionVar_4 = \TransitionVar_1 
%    \end{aligned}
  }
  \infer1[(Expanding $\FlowEq$)]{
    \begin{array}{c}
      \Connected(A', \Filter) \land
      \FlowEq(A', \Filter) \land \mbox{}\\
      l_A = x_1 + x_2 + x_3 + x_4 + x_5 \land
      i = x_5 \land 
      n = x_1 + x_3 + x_4 \land \mbox{}\\
      n > i 
    \land l_A - i - n > 0
    \land i > 0
    \end{array}
%    \begin{aligned}
%    i = \TransitionVar_5 
%    \land l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k 
%    \land n > i > 0
%    \land l_A - i - n > 0
%    \land n = \TransitionVar_4 + \TransitionVar_1\\
%    \land \Connected(\mathcal{A}', \Filter) 
%    \land \FlowEq(\mathcal{A}', \Filter)
%    \end{aligned}
  }
  \infer1[(Expanding sum)]{
    \begin{array}{c}
      \Connected(A', \Filter) \land
      \FlowEq(A', \Filter) \land
      \left[l_A, i, n\right] =
      \sum\limits_{\Transition \in \Transitions_{\mathcal{A}'} } \Map(\Transition) \cdot \Filter(\Transition) \land\mbox{}\\
     n > i 
    \land l_A - i - n > 0
    \land i > 0
    \end{array}
  }
  \infer1[\Expand{}]{
    \Image{}_{\mathcal{A}', \Map}(\Filter, \left[l_A, i, n\right]) 
    \land n > i 
    \land l_A - i - n > 0
    \land i > 0
    }
\end{prooftree}
\caption{A proof tree for
\cref{const:s1-in-c-dd,const:s1-substring,const:something-before-and-after,const:more-inside-than-before}
from \cref{ex:string-constraints}, corresponding to handling the Parikh
automaton $\mathcal{A}'$ of \cref{fig:aut_a}.}\label{fig:derivation:single}
\end{figure}

\subsection{Correctness of \Calculus{}}\label{sec:single:correct}

Our correctness proof of \Calculus{} consists of two main parts: first, we show
that the construction of a proof always terminates, and then that each of the
proof rules in \cref{tbl:rules:single} is an equivalence transformation, i.e.,
does not change the set of satisfying assignments of a formula. In combination,
those two results immediately imply that \Calculus{} gives rise to a decision
procedure.

\subsubsection{\Calculus{} terminates}
\begin{lemma}\label{lma:single-terminates}
  Suppose $\SomeClause{}$ is a set of formulas in which the predicates
  $\Image$ and $\Connected$ only occur positively. There is no
  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
  $P_i$ by applying one of the rules in \cref{tbl:rules:single}.
\end{lemma}

\begin{proof}
  The rule~\Expand{} can only be applied finitely often since each
  application removes one $\Image$ predicate, and none of the rules
  introduce new instances of the predicate. The rule~\Subsume{} can
  only be applied finitely often since it strictly decreases the
  combined number of $\Image$ and $\Connected$ predicates in sets of
  formulas, and none of the rules increases that number.

  To show termination of \Split{} and \Propagate{}, observe that the
  $\Filter$ in a predicate~$\Connected(\Automaton, \Filter)$ is never
  updated on a proof branch, which means that the set of terms
  $\Filter(t)$ for $t \in \Automaton$ on every branch is finite. Each
  application of \Split{} and \Propagate{} adds a new
  formula~$\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$
  to a proof goal, which can only happen finitely often.
\end{proof}

\subsubsection{The rules in \cref{tbl:rules:single} are solution-preserving}

\begin{lemma}\label{lma:single-correct}
  Consider an application of one of the rules in
  \cref{tbl:rules:single}, with
  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
  conclusion~$\SomeClause$. An assignment~$\beta$ satisfies the
  conclusion~$\SomeClause$ if and only if it satisfies one of the
  premises~$\SomeClause_i$.
\end{lemma}

\begin{proof}
  This property has to be shown by analysing the possible applications
  of each proof rule.

  \Expand{} unfolds the definition of the $\Image$ predicate. To show
  that the rule is solution-preserving, we prove the equivalence of the
  upper and lower sets of formulas:
  \begin{itemize}
  \item Assume that $\beta$ satisfies the conclusion, which means that
    there is some accepting path
    $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ with
    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
    \Path)$ and $\val_\beta(\MonoidElement) =
    \Map(\WordOf(\Path))$. Since immediately implies that $\beta$
    satisfies $\Connected(\Automaton, \Filter)$, since a path is
    connected, and $\FlowEq(\Automaton, \Filter)$ since an accepting
    path satisfies the flow equations. The
    equation~$\MonoidElement = \sum_{\Transition \in
      \Transitions_\Automaton}\Filter(\Transition) \cdot
    \Map(\Transition)$ holds because of
    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
    \Path)$.
  \item Assume that $\beta$ satisfies the premise, which implies that
    $\val_\beta(\Filter)$ describes a consistent, connected flow of
    the automaton. By the same argument as in
    \cite{generate-parikh-image}, this flow
    can be mapped to an accepting path~$\Path$ of $\Automaton$ such
    that each transition~$\Transition$ occurs on $\Path$ exactly
    $\val_\beta(\Filter(\Transition))$ times. Together with the equation
    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
    \Path)$, this implies that $\beta$ satisfies $\SinglePredicateInstance$.
  \end{itemize}

  In \Split{}, we make use of the fact that $\Filter(\Transition)$ is
  $\Naturals$-valued by definition. For any $\beta$, clearly exactly
  one of $\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$ will
  be satisfied, implying the property.

  For \Propagate{}, suppose that $\Dominates(C, \Automaton, \Transition)$, which
  means that every accepting path containing~$\Transition$ contains at least one
  of the transitions in $C$. For a $\beta$ satisfying $\Connected(\Automaton,
  \Filter)$. This means that every accepting path $p$ where $\Transition \in p$
  has at least one transition $\Transition' \in C$ such that
  $\Filter(\Transition') = 0$. This means that also
  $\val_\beta(\Filter(\Transition)) = 0$ has to hold since no unbroken path
  containing $\Transition$ exists.

  Finally, for \Subsume{}, observe that if \Split{} cannot be applied,
  then a goal must contain~$\Filter(\Transition) = 0$ or
  $\Filter(\Transition) > 0$ for every $\Transition$. In case the
  formulas in $\SomeInequalities$ are inconsistent, an application of
  \Subsume{} is trivially solution-preserving; therefore assume that
  $\SomeInequalities$ is consistent, which means that it contains
  exactly one of $\Filter(\Transition) = 0$ or
  $\Filter(\Transition) > 0$ for each $\Transition$. Since
  \Propagate{} is not applicable, the transitions~$\Transition$ with
  $\Filter(\Transition) > 0$ must form a connect sub-graph of the
  automaton; this means that $\Connected(\Automaton, \Filter)$ is
  redundant as it is implied by $\SomeInequalities$.
\end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: