### 2023-07-12 multiple.tex
tried to fix Materialise

We now generalise our calculus to natively work with intersections of
regular languages, or equivalently products of automata. For this
extension, we change the main predicate~$\Image$ to be indexed by a
vector of automata~$\Tuple{\Automaton_1, \ldots, \Automaton_k}$.  For
simplicity, we assume that the sets of states of the $k$ automata (and
therefore also the transition sets) are pairwise disjoint. 
\begin{definition}\label{def:multiple}
  Suppose $\Automaton_1, \ldots, \Automaton_k$ are automata,
  $\Map : \Alphabet^* \to \Monoid$ is a map to a commutative monoid
  $\Monoid$,
  $\Filter: \bigcup_{i=1}^k \Transitions_{\Automaton_i} \to \Naturals$
  is a transition selection function, and
  $\MonoidElement \in \Monoid$. The predicate
  $\ImagePredicate{\Tuple{\Automaton_1, \ldots,
      \Automaton_k}}{\Map}{\Filter}{\MonoidElement}$ is true exactly
  when there are accepting paths~$\Path_1, \ldots, \Path_k$ of the
  respective automata, such that for each $i \in \{1, \ldots, k\}$ and
  $\Transition \in \Transitions_{\Automaton_i}$ it holds that
  \begin{itemize}
  \item the multiplicity of $\Transition$ on $\Path_i$ is consistent with
    $\Filter$, that is,
    $\Filter(\Transition) = \TransitionCount(\Transition, \Path_i)$,
  \item the automata all accept the same
    word~$\WordOf(\Path_i) = \WordOf(\Path_1)$, and
  \item the accepted word is mapped to
    $\MonoidElement = \Map(\WordOf(\Path_i)) =
    \Map(\WordOf(\Path_1))$.
  \end{itemize}
\end{definition}

  \begin{table}[t]
    \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
      Name & \text{Rule} & Side conditions\\
      \midrule

      % EXPAND
      \ExpandM & 
      \inferrule
      {
       {\begin{array}{c}
           \left\{
           \FlowEq(\Automaton_i, \Filter),~~
           \Connected(\Automaton_i, \Filter),~~
           \MonoidElement = \sum_{\Transition \in \Transitions_{\Automaton_i}} \Filter(\Transition) \cdot \Map(\Transition)
           \right\}_{i=1}^k,
           \\[4ex]
          \ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, ~
          \SomeInequalities, \SomeClause
        \end{array}}
        }
      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} & 
      None \\[5ex]
      \Materialise &
      \inferrule
      {{
         \begin{array}{c}
           \FlowEq(\Automaton', \Filter'),~
           \Connected(\Automaton', \Filter'),~
           \BindingSum(\Automaton_i, \Automaton_j, \Filter')
           \\\ImagePredicate{
                     \Tuple{\Automaton_1, \Automaton_{i-1}, \Automaton_{i+1}, \ldots,
                     \Automaton_{j-1}, \Automaton_{j+1}, \ldots, \Automaton_k,
                     \Automaton'}
           }{\Map}{\Filter'}{\MonoidElement},~ \SomeInequalities, \SomeClause
           \end{array}}
           }
      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_i, \ldots, \Automaton_j, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement},~ \SomeInequalities, \SomeClause} &
      $\begin{array}{@{}l@{}}
         \Filter' = \Extend(\Automaton',\Filter),\\
         1 \leq i < j \leq k, \\
      \Automaton' =
         \Filter_{\SomeInequalities}(\Automaton_i) \times \Filter_{\SomeInequalities}(\Automaton_j)\\
      \end{array}$ \\
      \bottomrule
      \end{tabular}
      \caption{Additional derivation rules for products of arbitrarily many automata.}\label{tbl:rules:multi}
    \end{table}

    For the calculus (\cref{tbl:rules:multi}), we first extend
    $\Expand$ to generate flow equations and instances of $\Connected$
    for each automaton, resulting in a new rule $\ExpandM$.  Unlike
    \Expand{}, \ExpandM{} does not remove the $\Image$~predicate,
    since it is needed to keep track of the currently considered
    partial products.

The rule~$\Materialise$ introduces the product of two individual
automata~$\Automaton_i, \Automaton_j$; this step eliminates
$\Automaton_i, \Automaton_j$ as index of the $\Image$ predicate, and
instead adds the product of the automata restricted to the
transitions that can still be taken. The rule
also introduces the flow equations and the $\Connected$ predicate.

In $\Materialise$, we use the following notation for pruning away
parts of an automaton based on the transition selection
function~$\sigma$, only keeping those transitions for which~$\Filter$
is positive:
\begin{equation*}
  \Filter_{\SomeInequalities}(\Automaton) =
  \Tuple{\States, \InitialState, \AcceptingStates,
\Set{\Transition \in \Transitions \SuchThat (\Filter(\Transition) = 0) \not\in E}}
\end{equation*}
This filtering operation can be optimised to also eliminate states from
$Q$ and $F$ that become unreachable; this is kept implicit for sake of
presentation at this point.

The rule~$\Materialise$ has to connect the newly introduced
product to the previous automata
$\Automaton_i, \Automaton_j$. This is done by extending the
selection function~$\Filter$ to $\Filter'$, mapping the transitions of the product
to fresh variables~$x_{\Transition}$:
\begin{equation*}
  \Extend(\Automaton',\sigma)(t)
  ~~=~~
  \begin{cases}
    x_{\Transition} & t \in \Transitions_{\Automaton'}
    \\
    \Filter(t) & \text{otherwise}
  \end{cases}
\end{equation*}

The multiplicity of transitions in the product then has to be related
to the multiplicities in the individual automata, modelled using the
$\BindingSum$ predicate. The predicate expresses that the multiplicity
of a transition~$\Transition \in \Transitions_{\Automaton_i}$ in
$\Automaton_i$ has to coincide with the sum of the multiplicities of
transitions in $\Filter_{\SomeInequalities}(\Automaton_i) \times \Filter_{\SomeInequalities}(\Automaton_j)$ derived from $t$,
and similarly for $\Automaton_j$:
%
  $$
  \BindingSum(\Automaton_i, \Automaton_j, \Filter') ~~=~~
  \begin{aligned}
  & \Set{ 
    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State'_R}}} \Filter'(\Transition')
  \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_i} } \cup\mbox{} \\ 
  & \Set{
    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State'_L, \State'}}} \Filter'(\Transition') \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_j}
  }
  \end{aligned}
$$

For instances of precisely one automaton, neither rule applies, and we perform
the calculus as before.

\subsection{An Example}\label{sec:multiple:example}

We return to our example in \cref{sec:intuition}. We also reuse the
definition of $\Map$ as the simple projection to extract the counter increments,
so in this case, $\Map(\Transition) = \left[l_{B+}\right]$ for every
transition~$\Transition$ since $\mathcal{B}$ only counts length. The reader is
advised to review \cref{fig:aut_a_annotated,fig:aut_b_annotated} from
\cref{sec:intuition} while going through this example.
The transition variables of the figures match the ones used here.

We extend \cref{eq:single-filter} from \cref{sec:single:example} with
definitions for $B$ shown in \cref{eq:multi-filter}:
\begin{equation}\label{eq:multi-filter}
    \begin{aligned}
        & \Filter(\FromLabelTo{A}{a / l_{B+}}{F_B} ) = \TransitionVar_6 
        & \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) = \TransitionVar_7 
        & \Filter(\FromLabelTo{B}{b / l_{B+}}{I_B}) = \TransitionVar_8 
        &\Filter(\FromLabelTo{B}{c / l_{B+}}{B}) = \TransitionVar_9  \\
        & \Filter(\FromLabelTo{B}{c / l_{B+}}{F_B})= \TransitionVar_{10}
        & \Filter(\FromLabelTo{I_B}{a / l_{B+}}{A})= \TransitionVar_{11}
        & \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B})= \TransitionVar_{12}
        \FlowEq(\mathcal{B}, \Filter) &=  \begin{aligned}
            1 + \TransitionVar_{8}   &= \TransitionVar_{11} + \TransitionVar_{12}\\
            \land \TransitionVar_{11} &= \TransitionVar_6\\
            \land \TransitionVar_{12} &= \TransitionVar_8 + \TransitionVar_{10}
        \end{aligned}
    \end{aligned}
\end{equation}

The only possible rule at the start is $\ExpandM$, which we use to add the
corresponding constraints on each automaton of the product as we would have had
in the single-automaton version. After that, we expand the various sums and apply
light equality propagation.

We then continue by repeating the steps of \cref{sec:single:example} shown in
\cref{fig:derivation:single}, since the steps to propagate
\cref{ex:string-constraints} across the automaton $\mathcal{A}'$ still apply for the
product. Note that this fixes the values of all transitions but one of $\mathcal{A}'$ as
constants, including the free counter variables $i, n$, whose new constant
values we propagate everywhere.

Once we have performed reasoning by equality to eliminate transition variables,
we obtain the version of $\mathcal{B}$ shown in \cref{fig:aut_b_annotated} and represented
by the clauses in the final node in the tree. This corresponds to the end of
\cref{sec:intuition:algebra}. Our only options now are to either split on which
path we take from $I_B$ by case-splitting on~$\TransitionVar_{12}$ or to
directly invoke \Materialise{}. The latter would produce a shorter tree in this
instance but might lead to a larger product being computed, so we (and indeed
\Catra{}) pick \Split{}. This leads to an opportunity for propagation since
$\TransitionVar_{12} = 0$ cuts off $\TransitionVar_{9}$ from $I_B$ on the left
branch. Note that the same opportunity does not present itself on the right
branch, though; the use of the upper path to state~$A$ does not preclude visiting
state~$B$.

In both cases, we have comparatively simple automata left, so we apply
\Materialise{} to obtain their product, which for both branches is empty. An
empty product will have no outgoing transitions from its initial state, and so
will produce a flow equation $1 = 0$ since the right-hand balancing side will
always be the empty sum. We can then close the proof and know that the problem
is unsatisfiable.

\begin{figure}
  \centering
\begin{prooftree}
  {% % LEFT TREE: here x_12 is zero
  \hypo{
    \begin{array}{l}
      \ldots \boxed{\land 1 = 0} \\
      \land \Image{}_{\emptyset, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A})
    \end{array}
  }
  \infer1[\Materialise]{
    \begin{array}{llll}
      \TransitionVar_9 = 0 
      \land \TransitionVar_7 = l_A - 2 \\
      \land \TransitionVar_2 = l_A - 3
      \land \TransitionVar_{8} = 0 \\
      \land \TransitionVar_6 = \TransitionVar_{11} = 1
      \land \TransitionVar_{10} = 0 \\
      \land \Connected(B, \Filter)\\
      \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A})
    \end{array}
  }
  \infer1[\Propagate{}]{
    \begin{array}{llll}
      \TransitionVar_{12} = 0 
      \land \TransitionVar_7 = l_A - \TransitionVar_9 - 2 \\
      \land \TransitionVar_2 = l_A - 3 
      \land \TransitionVar_{8} = 0 \\
      \land \TransitionVar_6 = \TransitionVar_{11} = 1 \\
      \land \TransitionVar_{10} = 0 
      \land \Connected(\mathcal{B}, \Filter) \\
      \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
    \end{array}
    }
  }% END LEFT TREE
  {% RIGHT TREE % here x_12 is > 0
  \hypo{
    \begin{array}{l}
      \ldots \boxed{\land 1 = 0} \\
      \land \Image{}_{\emptyset, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A})
    \end{array}
  }
  \infer1[\Materialise]{
    \begin{array}{rrrr}
      \TransitionVar_{12} > 0 \land \\
      \TransitionVar_7 = l_A - \TransitionVar_9 - 2\TransitionVar_{11}-2\TransitionVar_{12} \land \\
      \TransitionVar_2 = l_A - 3 \land \\
      \TransitionVar_{8} = \TransitionVar_{11} + \TransitionVar_{12} - 1 \land \\
      \TransitionVar_6 = \TransitionVar_{11} \land\\
      \TransitionVar_{10} = 1 - \TransitionVar_{11} \land\\
      \Connected(\mathcal{B}, \Filter) \land\\
      \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
    \end{array}
  }
  }% END RIGHT TREE
  \infer2[\Split{} $\TransitionVar_{12} = 0$]{
  \begin{array}{llll}
    \TransitionVar_7 = l_A - \TransitionVar_9 - 2\TransitionVar_{11}-2\TransitionVar_{12} \\
    \land \TransitionVar_2 = l_A - 3 \\
    \land \TransitionVar_{8} = \TransitionVar_{11} + \TransitionVar_{12} - 1 \\
    \land \TransitionVar_6 = \TransitionVar_{11} \\
    \land \TransitionVar_{10} = 1 - \TransitionVar_{11} \\
    \land \Connected(\mathcal{B}, \Filter)\\
    \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
  \end{array}
  }
  \infer1[Expand $\sum$, \EquationReasoning] {
    \begin{array}{llll}
      l_A = \sum_{k = 6}^{k = 12} \TransitionVar_k
      \land \TransitionVar_2 = l_A - 3
      \land 1 + \TransitionVar_{8}   = \TransitionVar_{11} + \TransitionVar_{12}
      \land \TransitionVar_{11} = \TransitionVar_6 \\
      \land \TransitionVar_{12} = \TransitionVar_8 + \TransitionVar_{10}
      \land \Connected(\mathcal{B}, \Filter) 
      \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
    \end{array}
  }
  \infer1[\EquationReasoning: expand $\FlowEq$]{
    \begin{array}{llll}
      l_A = \sum_{k = 6}^{k = 12} \TransitionVar_k
      \land \TransitionVar_2 = l_A - 3
      \land \FlowEq(\mathcal{B}, \Filter) \\
      \land \Connected(\mathcal{B}, \Filter) 
      \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
    \end{array}
  }
  \infer1[Repeat \cref{fig:derivation:single}, $l_A = l_B$.]{
    \begin{array}{llll}
      l_B = \sum_{k = 6}^{k = 12} \TransitionVar_k
      \land i = \TransitionVar_5
      \land l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k \\
      \land n = \TransitionVar_4 + \TransitionVar_1
      \land \FlowEq(\mathcal{B}, \Filter)
      \land \Connected(\mathcal{A}', \Filter)\\
      \land \Connected(\mathcal{B}, \Filter)
      \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B}) \\
      \land l_A = l_B
      \land l_A - i - n > 0
      \land 0 < i < n
    \end{array}
  }
  \infer1[\ExpandM]{
    \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B})
    \land l_A = l_B % Implied constraint
    \land l_A - i - n > 0
    \land 0 < i < n
    }
\end{prooftree}
\caption{A derivation for \Calculus{} on the Parikh image strings for the
constraints of \cref{ex:string-constraints}. Note the constant propagation
for~$n$!}\label{fig:derivation:multi}
\Description[]{}% This isn't a problem; it's all text
\end{figure}

\subsection{Correctness of \Calculus{} for Products of Automata}

Since \cref{tbl:rules:multi} only extends the existing rules of
\cref{tbl:rules:single}, we focus on the differences compared
to the calculus for a single automaton.

\subsubsection{\Calculus{} for products of automata terminates}
\begin{lemma}\label{lma:multi-terminates}
  Suppose $\SomeClause{}$ is a set of formulas in which the product
  version of $\Image$ only occurs positively. There is no
  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
  $P_i$ by applying one of the rules in \cref{tbl:rules:multi}.
\end{lemma}

\begin{proof}
  The rule \Materialise{} can similarly only be used finitely many times, as
  each application reduces the number of automata in the product of $\Image$ by
  one automaton, until only one remains and \cref{lma:single-terminates} for
  single-automaton instances apply.

  This implies that also the rule~\ExpandM{} can only be applied
  finitely often since its side condition only allows applying it once
  per instance of an $\Image$ predicate containing a product, and only finitely
  many instances of $\Image$ can be introduced on each branch.
\end{proof}

\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}

Since our calculus now includes a rule introducing new variables, the
\Materialise{} rule, we have to slightly generalise the notion of
solution-preservation:
%
\begin{lemma}\label{lma:multi-correct}
  Consider an application of one of the rules in
  \cref{tbl:rules:multi}, with
  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
  conclusion~$\SomeClause$. An assignment~$\beta$ (over the symbols in
  $\SomeClause$) satisfies the conclusion~$\SomeClause$ if and only if
  there is an extension~$\beta' of \beta$ satisfying one of the
  premises~$\SomeClause_i$.
\end{lemma}

\begin{proof}
  We have to consider the two new rules in \cref{tbl:rules:multi}. The
  result is immediate for \ExpandM{}, since this rule does not remove the
  $\Image$ predicate from a proof goal, and the newly introduced formulas
  are all implied by the $\Image$ predicate.

  For \Materialise{}, observe that the existence of an accepting path
  in $\Automaton_i \times \Automaton_j$ is equivalent to the existence
  of individual paths in $\Automaton_i, \Automaton_j$ accepting the
  same word. The path in the product will satisfy the flow equations
  and connectedness, and it will be related to the individual paths as
  stipulated by the \BindingSum{} predicate.
\end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: