### 2024-02-22 appendix.tex
Flesh out evaluation somewhat, move figures from appendix
Closes: #73
Closes: #74

Also addresses: #78

\subsection{Example instance for \Catra{}}
\lstinputlisting[caption={An example input file for \Catra{} for \cref{ex:string-constraints} in \cref{sec:intuition}, illustrating every major syntax element. From beginning to end: synchronised (product) automata using the keyword \texttt{synchronised} (automata A and B), labels (except those with ranges), register increments, and constraints on the final values of their counters.}, label=lst:input-example]{example.par}

%\subsection{SMT-LIB encoding of the string constraint example}
%\lstinputlisting[caption={An SMT-LIB encoding of \cref{ex:string-constraints}.}, label=lst:string-constraints]{example.smt2}