### 2024-03-06 data.tex
Data statement

A reproduction package featuring all relevant logs is available~\cref{artefact}.
Additionally, \Ostrich{} and \Catra{} are both available as living software
on GitHub at \url{https://github.com/amandasystems/catra} and \url{https://github.com/uuverifiers/ostrich}
respectively. A version of \Ostrich{} modified to use CA-Str is available in the \texttt{cea-catra}~branch
and comes bundled with the reproduction package.

All source code to generate the artefact and the paper itself is \emph{also} available on 
GitHub at~\url{https://github.com/amandasystems/oopsla-artefact}.

\Ostrich{} and \Catra{} are both open source under a 3-clause BSD~license.