### 2024-03-08 data.tex
Add artefact, update evaluation

A reproduction package featuring all relevant logs and all benchmarked instances is available~\cite{artefact}.
\Catra{} and \Ostrich{} are both available as living software
under a 3-clause BSD~license on GitHub at \url{https://github.com/amandasystems/catra} and \url{https://github.com/uuverifiers/ostrich}
respectively. A version of \Ostrich{} modified to use CA-Str is available in the \texttt{cea-catra}~branch. All source code to generate the paper itself
is available at~\url{https://github.com/amandasystems/oopsla-artefact}.